### 2022-06-10 acml/acml22_paper.tex
started acml paper

%\documentclass[wcp,gray]{jmlr} % test grayscale version
\documentclass[wcp]{jmlr}

% givasile packages
\usepackage{bbm}
\usepackage{xfrac}

% givasile commands
\newcommand{\xc}{\mathbf{x}_c}
\newcommand{\Xc}{\mathbf{X}_c}
\newcommand{\X}{\mathbf{X}}
\newcommand{\Ysloc}{Y_{s}^{\text{local}}}
\newcommand{\ysloc}{y_{s}^{\text{local}}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\xii}{\mathbf{x}^{(i)}}
\newcommand{\yi}{y^{(i)}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\1}{\mathbbm{1}}
\newcommand{\D}{\mathcal{D}}

% The following packages will be automatically loaded:
% amsmath, amssymb, natbib, graphicx, url, algorithm2e

%\usepackage{rotating}% for sideways figures and tables
\usepackage{longtable}% for long tables

% The booktabs package is used by this sample document
% (it provides \toprule, \midrule and \bottomrule).
% Remove the next line if you don't require it.
\usepackage{booktabs}
% The siunitx package is used by this sample document
% to align numbers in a column by their decimal point.
% Remove the next line if you don't require it.
%\usepackage[load-configurations=version-1]{siunitx} % newer version
%\usepackage{siunitx}
%\usepackage{natbib}

% Do not comment the following commands:
\pagenumbering{gobble}
\newcommand{\cs}[1]{\texttt{\char`\\#1}}
\makeatletter
\let\Ginclude@graphics\@org@Ginclude@graphics 
\makeatother

\jmlrvolume{}
\jmlryear{2022}
\jmlrworkshop{ACML 2022}

\title[Short Title]{Full Title of Article}

 % Use \Name{Author Name} to specify the name.
 % If the surname contains spaces, enclose the surname
 % in braces, e.g. \Name{John {Smith Jones}} similarly
 % if the name has a "von" part, e.g \Name{Jane {de Winter}}.
 % If the first letter in the forenames is a diacritic
 % enclose the diacritic in braces, e.g. \Name{{\'E}louise Smith}

 % Two authors with the same address
 % \author{\Name{Author Name1} \Email{abc@sample.com}\and
 %  \Name{Author Name2} \Email{xyz@sample.com}\\
 %  \addr Address}

 % Three or more authors with the same address:
 % \author{\Name{Author Name1} \Email{an1@sample.com}\\
 %  \Name{Author Name2} \Email{an2@sample.com}\\
 %  \Name{Author Name3} \Email{an3@sample.com}\\
 %  \Name{Author Name4} \Email{an4@sample.com}\\
 %  \Name{Author Name5} \Email{an5@sample.com}\\
 %  \Name{Author Name6} \Email{an6@sample.com}\\
 %  \Name{Author Name7} \Email{an7@sample.com}\\
 %  \Name{Author Name8} \Email{an8@sample.com}\\
 %  \Name{Author Name9} \Email{an9@sample.com}\\
 %  \Name{Author Name10} \Email{an10@sample.com}\\
 %  \Name{Author Name11} \Email{an11@sample.com}\\
 %  \Name{Author Name12} \Email{an12@sample.com}\\
 %  \Name{Author Name13} \Email{an13@sample.com}\\
 %  \Name{Author Name14} \Email{an14@sample.com}\\
 %  \addr Address}

 % Authors with different addresses:
  \author{\Name{Author Name1} \Email{abc@sample.com}\\
  \addr Address 1
  \AND
  \Name{Author Name2} \Email{xyz@sample.com}\\
  \addr Address 2
 }

\editors{Emtiyaz Khan and Mehmet Gonen}

\begin{document}

\maketitle

\begin{abstract}
This is the abstract for this article.
\end{abstract}
\begin{keywords}
List of keywords separated by semicolon.
\end{keywords}

\section{Introduction}
Main contents here.

\section{Related Work}

\section{Probabilistic ALE plots}

\subsection{The ALE method}\label{sec:ALE}

In this section, we introduce the reader to the feature effect method
ALE. Given that \(f: \R^S \rightarrow \R \) is known, we can measure
the \textit{local} effect of the \(s\)-th feature at a specific point
\( \x = (\xc, x_s)\) of the input space \(\mathcal{X}\), as
\(f_s(\x) = \sfrac{\partial f(\x)}{\partial x_s}\). ALE models the
\textit{local} feature effect at \(x_s\) as an expectation over the
distribution of the unkown (latent) features
\(\E_{p(\Xc;x_s)}[f_s(x_s, \xc)]\). Afterwards, ALE measures the
\textit{global} effect at \(x_s\) as an accumulation of the expected
\textit{local} effects:
%
\begin{equation}
  \label{eq:ale-definition}
  f_{\mathtt{ALE}}(x_s) =
  \int_{x_{s, \text{min}}}^{x_s} \E_{p(\Xc;x_s=z)}[f_s(x_s, \xc)] \partial z
\end{equation}

In real cases, it is infeasible to compute
eq.~\eqref{eq:ale-definition} analytically. Therefore, we reside on
estimating the effect from the training set. Let's denote the
available dataset as \(\D = \{\xii, \yi \}_{i=1}^N\), where \( \xii \)
is the \(i\)-th feature vector of the training data and \(\yi\) is the
\(i\)-th label. Apley et. al proposed spliting the axis into \( K \)
equal-sized bins, find the set of points that lie in each bin, i.e.
\( \mathcal{S}_k = \{ \xii : x_s^{(i)} \in [z_{k-1}, z_k) \} \) and,
finally, find the local effect at each bin as the mean value of the
population, i.e.
\(\hat{\mu}_{s,k} = \frac{1}{|\mathcal{S}_k|} \sum_{i:\xii \in
  \mathcal{S}_k} [f_s(\xii)]\). The global effect at \(x_s\) is then
estimated through:
%
\begin{equation}
  \label{eq:ale-approximation}
  \hat{f}_{\mathtt{ALE}}^s(x)
  = \Delta x \sum_{k=1}^{k_{x_s}} \hat{\mu}_{s,k}
  = \Delta x \sum_{k=1}^{k_{x_s}} \frac{1}{|\mathcal{S}_k|} \sum_{i:\xii \in \mathcal{S}_k} [f_s(\xii)]
\end{equation}
%
where \(\Delta x\) is the bin size and \(k_{x_s}\) the bin-index of the value \(x_s\).
%
%
\subsection{ALE plots with uncertainty quantification}

ALE directly defines the local effect as an expectation over the
unknown features and the global effect as integral over the
expectations. Therefore, both local and global effects are modeled as
simple variables. In constrast, for integrating ALE plots in a
probabilistic framework, we model the local effect as a random
variable, dependent on the random variable \(\Xc\) representing the
latent features:

\begin{equation}
    \Ysloc \sim p(\ysloc ; x_s)
    = \int_{\xc} \delta(\ysloc-f_s(\x)) p(\xc;x_s) \partial \xc
    \label{eq:local-effect-rv}
\end{equation}
%
Following the core idea of ALE, we define the global feature effect as
the random variable \(Y_s\) which is the integration of the local
effects:

\begin{equation}
  Y_s \sim p(y_s ; x_s) = \int_{x_{s,min}}^{x_s} p(\ysloc;x_s=z) \partial z
  \label{eq:global-effect-rv}
\end{equation}
%
Through eqs.~(\ref{eq:local-effect-rv}), (\ref{eq:global-effect-rv})
we formulate the ALE in a fully probabilistic manner. We notice, that
the definition of ALE as given by~\cite{Apley2020} is simply the
expected value of \(Y_s\), i.e.
\(f_{\mathtt{ALE}}(x_s) = \E[Y_s ; x_s]\). For simpler notation we
denote \(\E[Y_s ; x_s]\) as \(\mu_s (x_s)\).

For measuring the uncertainty, it is also important to capture the
variance of the global effect. Notating the variance of the feature
effect \(\var[Y_s ; x_s]\) as \(\sigma_s^2\), we can compute it
through:

\begin{equation}
  \sigma_s^2 (x_s)
  = \int_{x_{s,min}}^{x_s} \int_{\xc} p(\xc|x_s) (f_s(\x) - \mu_s(x_s))^2 \partial \xc (\partial z)^2
\end{equation}

As before, it is infeasible to compute the the variance
\(\sigma_s^2(x_s)\) in closed-form. Therefore, we reside on estimating
them from the available examples of the training set. The estimation
of the variance can also done through the available points:

\begin{equation}
  \label{eq:effect-var-approx}
  \hat{\sigma}_s^2(x_s)
  = (\Delta x)^2 \sum_{k=1}^{k_{x_s}} \hat{\sigma}_{s,k}^2
  = (\Delta x)^2 \sum_{k=1}^{k_{x_s}} \frac{1}{|\mathcal{S}_k|} \sum_{i:\xii \in \mathcal{S}_k} (f_s(\xii) - \hat{\mu}_{s,k})^2 
\end{equation}

\subsection{Unsupervised metric for assessing the quality of ALE plots}

\subsection{ALE plots with variable-size bins}

\subsection{Variable-size bins as an optimization problem}

\section{Experiments}

\subsection{Synthetic Data sets}

A figure in Fig.~\ref{fig:spiral}. Please use high quality graphics
for your camera-ready submission -- if you can use a vector graphics
format such as \texttt{.eps} or \texttt{.pdf}.
\begin{figure}[htp]
\begin{center}
\includegraphics[width=0.5\textwidth]{spiral.eps}
\caption{A spiral.}\label{fig:spiral}
\end{center}
\end{figure}

An example of citation~\cite{DBLP:conf/acml/2009}.

\subsection{Real Data sets}

\section{Conclusion}

\subsection{Subsection Title}

%\acks{Acknowledgements should go at the end, before appendices and references. You can uncomment this for the camera-ready version on paper acceptance.}

%\bibliographystyle{plain}
\bibliography{acml22}

\appendix

\section{Derivations}\label{apd:first}

Derivations

\begin{multline}
    \Ysloc \sim p(\ysloc ; x_s)
    = \int{} p(\ysloc|\xc;x_s) p(\xc;x_s) \partial \xc \\
    = \int_{\xc} \delta(\ysloc-f_s(\x)) p(\xc;x_s) \partial \xc
    = \int_{\xc} \1(\ysloc=f_s(\x)) p(\xc;x_s) \partial \xc
\end{multline}

Some statistics for the local feature effect:

\begin{equation}
  \E[ \Ysloc ; x_s] = \int_{\xc} p(\xc|x_s) f_s(\x) \partial \xc
\end{equation}

\begin{equation}
  \var[ \Ysloc ; x_s] = \int_{\xc} p(\xc|x_s) (f_s(\x) - \E[ \Ysloc ; x_s])^2 \partial \xc
\end{equation}

Some statistics for the global feature effect:

\begin{equation}
  \E[Y_s ; x_s] = \int_{x_{s,min}}^{x_s} \E(\ysloc;z) \partial z
\end{equation}

\begin{equation}
  \var[Y_s ; x_s] = \int_{x_{s,min}}^{x_s} \var[ \Ysloc ; x_s] (\partial z)^2
\end{equation}

\section{Second Appendix}\label{apd:second}

This is the second appendix.

\end{document}