### 2022-10-10 AISTATS2023/paper.tex
review on simulation experiments

\documentclass[twoside]{article}

\usepackage{aistats2023}
% If your paper is accepted, change the options for the package
% aistats2023 as follows:
%
%\usepackage[accepted]{aistats2023}
%
% This option will print headings for the title of your paper and
% headings for the authors names, plus a copyright note at the end of
% the first column of the first page.

% If you set papersize explicitly, activate the following three lines:
%\special{papersize = 8.5in, 11in}
%\setlength{\pdfpageheight}{11in}
%\setlength{\pdfpagewidth}{8.5in}

% If you use natbib package, activate the following three lines:
\usepackage[round]{natbib}
\renewcommand{\bibname}{References}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}
\bibliographystyle{plainnat}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xfrac}
\usepackage{hyperref}
\usepackage{graphicx}
\graphicspath{ {./../examples/} }
% If you use BibTeX in apalike style, activate the following line:
%\bibliographystyle{apalike}

\usepackage{tikz}
\usetikzlibrary{matrix,positioning,arrows.meta,arrows,fit,backgrounds,decorations.pathreplacing}
\tikzset{ mymat/.style={ matrix of math nodes, text height=2.5ex, text
depth=0.75ex, text width=6.00ex, align=center, column
sep=-\pgflinewidth, nodes={minimum height=5.0ex} }, mymats/.style={
mymat, nodes={draw,fill=#1} }, mymat2/.style={ matrix of math nodes,
text height=1.0ex, text depth=0.0ex, minimum width=5ex, % text
width=7.00ex, align=center, column sep=-\pgflinewidth }, }
\usetikzlibrary{shapes.geometric, arrows, backgrounds, scopes}
\usepackage{pgfplots} \pgfplotsset{width=6.75cm, compat=newest}
\usepackage[utf8]{inputenc} \DeclareUnicodeCharacter{2212}{−}
\usepgfplotslibrary{groupplots,dateplot}
\usetikzlibrary{patterns,shapes.arrows}

\newcommand{\dfdx}{f^s}
\newcommand{\xc}{\mathbf{x}_c}
\newcommand{\DY}{\mathbf{\Delta Y}}
\newcommand{\xb}{\mathbf{x}}
\newcommand{\Xcb}{\mathbf{X}_c}
\newcommand{\Xb}{\mathcal{X}}
\newcommand{\todo}[1]{[\textcolor{red}{#1}]}

\begin{document}

% If your paper is accepted and the title of your paper is very long,
% the style will print as headings an error message. Use the following
% command to supply a shorter title of your paper so that it can be
% used as headings.
%
%\runningtitle{I use this title instead because the last one was very long}

% If your paper is accepted and the number of authors is large, the
% style will print as headings an error message. Use the following
% command to supply a shorter version of the authors names so that
% they can be used as headings (for example, use only the surnames)
%
%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}

\twocolumn[

\aistatstitle{Uncertainty-aware Accumulated Local Effects (UALE) for quantifying the heterogeneity of instance-level feature effects}

\aistatsauthor{ Author 1 \And Author 2 \And  Author 3 }

\aistatsaddress{ Institution 1 \And  Institution 2 \And Institution 3 } ]

\begin{abstract}

Accumulated Local Effects (ALE) is a popular approach to explainable AI that quantifies how a feature influences the decisions of a model taking into account feature correlations. However, in case of strong interactions between features, instance-level feature effects may deviate from the average curve provided by ALE. Therefore, it is crucial to quantify this deviation, namely, the uncertainty of the effect. In this work, we define Uncertainty-aware ALE (UALE) to quantify and visualize, on a single plot, both the average effect and the uncertainty. Furthermore, as in ALE, UALE's approximation requires partitioning the feature domain into non-overlapping intervals (bin-splitting). The average effect and the uncertainty are then computed from the instances that lie in each bin.  In this work, we formally prove that to achieve an unbiased approximation of the uncertainty in each bin, bin-splitting must follow specific constraints. Based on this, we propose a method for automatically finding the optimal intervals, balancing the trade-off between estimation bias and variance~\todo{check if this phrase explains well the concept.} We demonstrate through synthetic and real datasets (a) the advantages of modeling the uncertainty with UALE compared to other methods and (b) the importance of appropriate bin splitting for an accurate approximation of the average effect and uncertainty.
\end{abstract}

\section{INTRODUCTION}

Recently, Machine Learning (ML) has been adopted in critical domains,
such as healthcare and finance. In these areas, we need a combination
of accurate predictions along with meaningful explanations to support
them. For this reason, there is an increased interest in Explainable
AI (XAI) to understand the decision mechanism of complex black-box
models. XAI literature distinguishes between local and global
techniques~\citep{Molnar2020interpretable}. Local methods provide
instance-level explanations, i.e.~explain the prediction for a
specific instance, whereas global methods summarize the entire model
behavior.

Global methods create a global explanation by aggregating the
instance-level explanations into a single interpretable outcome,
usually a number or a plot. When strong interactions between features
exist, the instance-level explanations may significantly deviate from
the aggregated outcome~\citep{herbinger2022repid}, a phenomenon called
\emph{aggregation bias}~\citep{mehrabi2021survey}. Therefore, global
methods should quantify not only the aggregated (averaged) effect, but
also the level of heterogeneity of instance-level explanations. The
same holds for Feature Effect (FE)~\citep{Gromping2020MAEP}, a class
of global XAI methods that isolate the impact of a single feature on
the output.\footnote{FE methods also isolate the combined effect of a
  pair of features to the output. Combinations of more than two
  features are uncommon, since they are difficult to estimate and
  visualize.} For example, a feature with zero average effect may
indicate no effect on the output or a highly positive effect in some
cases and a highly negative effect in others. The ambiguity about the
correct interpretation can be resolved, if the FE method informs about
the heterogeneity of instance-level explanations, a quantity we refer
to as the \emph{uncertainty of the effect}.

There are three widely-used FE methods: \emph{Partial Dependence
  Plots} (PDP)\citep{friedman2001greedy}, \emph{Marginal Plots}
(MP)\citep{apley2020visualizing} and \emph{Aggregated Local Effects}
(ALE)\citep{apley2020visualizing}. ALE has been established as the
state-of-the-art for measuring the average effect, since PDP and MP
have been criticized~\citep{Gromping2020MAEP} to compute erroneous
effects when the input features are (highly) correlated. However, ALE
faces two limitations, the first concerns ALE definition and the
second ALE approximation. Regarding the definition, ALE does not
inform the user about the uncertainty. In contrast, in PDP the
heterogeneity of instance-level effects can be identified by exploring
the Individual Conditional Expectations
(ICE)~\citep{goldstein2015peeking}. Regarding the approximation, ALE
requires partitioning the feature domain into non-overlapping
intervals (bin-splitting) and estimate the effect inside each interval
(bin-effect) from the training set instances that fall inside. Finding
an appropriate sequence of bins is of particular importance, since
ALE's interpretation is meaningful only inside each
bin~\citep{molnar2022}. However, this crucial step has not raised the
appropriate attention. With the current
approaches~\citep{apley2020visualizing, gkolemis22}, the domain is
split \emph{blindly}, i.e.~without considering the characteristics of
the instance-level effects, in \(K\) equisized bins.

In this paper, we present Uncertainty-aware Accumulated Local Effects
(UALE), a method that visualizes in a single plot both the average
feature effect and the uncertainty, i.e.~the level of heterogeneity of
instance-level effects. The average effect is defined as in ALE and
the uncertainty inherits ALE's advantages, i.e.~takes into account the
correlations between features. We also propose an accurate UALE
approximation. We formally prove that for an unbiased estimation of
the uncertainty, the partitioning of the feature domain must respect
specific conditions. Using that, we set-up an optimization problem
where we search for a partitioning, a sequence of \emph{variable-size}
intervals, (a) with enough instances (accurate estimation) and (b)
that respect the above conditions (unbiased estimation). Finally, we
propose a method for automatically finding the optimal
partitioning. We apply our method in multiple synthetic and real world
datasets for evaluating its performance. The contributions of this
paper are:

\begin{itemize}
\item UALE, a feature effect method that quantifies both the average
  effect and the uncertainty, i.e.,~the heterogeneity of instance level
  effects.
\item A formal proof of the conditions the feature domain partitioning
  must fulfill for an unbiased approximation of the uncertainty.
\item An accurate UALE approximation by automatically partitioning
  into variable-size intervals that respect the above conditions.
\end{itemize}

The implementation of our method and the code for
reproducing all the experiments is provided in the submission and will
become publicly available upon acceptance.

\section{BACKGROUND AND RELATED WORK}

We here review the existing methods for modeling the average effect
and the heterogeneity of instance-level effects and we focus on ALE
definition~\citep{apley2020visualizing} and
approximation~\citep{apley2020visualizing, gkolemis22}.

\paragraph{Notation.} We refer to random variables (rv) using
uppercase \( X \), to simple variables with plain lowercase \( x \)
and to vectors with bold \( \xb \). Often, we partition the input
vector \(\xb \in \mathbb{R}^D\) to the feature of interest
\(x_s \in \mathbb{R} \) and the rest of the features
\(\xc \in \mathbb{R}^{D-1}\). For convenience we denote it as
\((x_s, \mathbf{x}_c)\), but we it refers the vector
\((x_1, \cdots , x_s, \cdots, x_D)\). Equivalently, we denote the
corresponding rv as \(\mathbf{X} = (X_s, \mathbf{X}_c)\). When we
refer to an instance of the training set, we use
\(\xb^i= (x_s^i, \xc^i) \).

% The black-box function is
% \(f : \mathbb{R}^D \rightarrow \mathbb{R}\) and the FE of the \(s\)-th
% feature is \(f^{\mathtt{<method>}}(x_s)\), with \(\mathtt{<method>}\)
% indicating the particular method in use.\footnote{An extensive list of
%   all symbols used in the paper is provided at the Appendix.}

\subsection{Feature Effect Methods}
\label{sec:feat-effect-meth}

The three well-known feature effect methods are: PDP, MP and ALE. PDP
formulates the FE of the \(s\)-th feature as an expectation over the
marginal distribution \(\mathbf{X}_c\), i.e.,
\(f^{\mathtt{PDP}}(x_s) =
\mathbb{E}_{\mathbf{X}_c}[f(x_s,\mathbf{X}_c)]\), whereas MP
formulates it as an expectation over the conditional
\(\mathbf{X}_c|X_s\), i.e.,
\(f^{\mathtt{MP}}(x_s) = \mathbb{E}_{\mathbf{X}_c|x_s}[f(x_s,
\mathbf{X}_c)]\). ALE defines the local effect at \(x_s = z\) as
\(f^s(z, \xc) = \frac{\partial f}{\partial x_s} (z, \xc)\). The local
explanations are then weighted by the conditional distribution
\(p(\xc|x_s=z)\) to compute the averaged effect \(\mu(z)\). ALE is the
accumulation of the averaged local effects:

\begin{equation}
  \label{eq:ALE}
  f^{\mathtt{ALE}}(x_s) = \int_{x_{s,min}}^{x_s} \underbrace{\mathbb{E}_{\Xcb|X_s=z}\left [f^s (z, \Xcb)\right ]}_{\mu(z)} \partial z
\end{equation}

Eq.~(\ref{eq:ALE}) has important advantages compared to PDP and MP in
cases of correlated features. In these cases, PDP integrates over
unrealistic instances due to the use of the marginal distribution
\(\mathbf{X}_c \), and MP computes aggregated effects, i.e., imputes
the combined effect of sets of features to a single feature. ALE
resolves both issues.

\subsection{Heterogeneity Of Instance-Level Effects}
\label{sec:quant-heter-effects}

FE methods show what is the average (global) effect on the output, if
the value of a specific feature is increased/decreased. For completing
the explanation about a feature, we need to also know to what extent
the local effects deviate from the global explanation, i.e.~to
quantify the uncertainty of the average effect.

ICE and d-ICE\citep{goldstein2015peeking} provide a set of curves that
are plot on top-of PDP. Both methods produce one curve for each
instance of the training set;
\(f^{\mathtt{ICE}}_i(x_s) = f(x_s, \xc^i)\) for ICE and
\(f^{\mathtt{d-ICE}}_i(x_s) = \frac{\partial f}{\partial x_s} (x_s,
\xc^i)\) for d-ICE. The user then visually observes if the curves are
homogeneous and to what extent they deviate from the PDP. Some methods
try to automate the aforementioned visual exploration, by grouping
(d-)ICE plots into clusters~\citep{molnar2020model,
  herbinger2022repid, britton2019vine}. Other approaches, like
H-Statistic\citep{friedman2008predictive}, Greenwell's interaction
index\citep{greenwell2018simple} or SHAP interaction
values\citep{lundberg2018consistent} inform indirectly about
heterogeneous effects, quantifying interactions between the input
features. A strong interaction index is an indicator for the existence
of heterogeneous effects.

The aforementioned approaches are under two limitations. They either
do not quantify the uncertainty of the FE directly or they are based
on PDPs, and, therefore, they are subject to the failure modes of PDPs
in cases of correlated features\citep{baniecki2021fooling}, as we will
show in-depth in Section~\ref{sec:simulation-examples-1}. To the best
of our knowledge, no work exist so far that quantifies the
heterogeneous effects based on the formulation of ALE.

\subsection{ALE Approximation}
\label{sec:ale-approximation}

In ML scenarios, ALE is estimated from the instances of the training
set.~\citep{apley2020visualizing} proposed dividing the feature domain
in \(K\) equisized bins and estimating the local effects in each bin
by evaluating the black box-function at the bin limits:
\begin{equation}
  \label{eq:ALE_accumulated_mean_est}
  \hat{f}^{\mathtt{ALE}}(x_s) = \sum_{k=1}^{k_x} \frac{1}{|\mathcal{S}_k|} \sum_{i:\mathbf{x}^i \in
    \mathcal{S}_k} \left [ f(z_{k}, \xc^i) - f(z_{k-1}, \xc^i)) \right ]
\end{equation}
We denote as \(k_x\) the index of the bin that \(x_s\) belongs to,
i.e. \(k_x: z_{k_x-1} \leq x_s < z_{k_x} \) and \(\mathcal{S}_k\) is
the set of training instances of the \(k\)-th bin, i.e.
\( \mathcal{S}_k = \{ \xb^i : z_{k-1} \leq x^i_s < z_{k} \}
\). Afterwards,~\citep{gkolemis22} proposed the Differential ALE
(DALE) that uses the partial derivatives:
\begin{equation}
  \label{eq:DALE_accumulated_mean_est}
  \hat{f}^{\mathtt{DALE}}(x_s) = \Delta x \sum_{k=1}^{k_x} \frac{1}{|\mathcal{S}_k|} \sum_{i:\mathbf{x}^i \in
    \mathcal{S}_k} f^s(\mathbf{x}^i)
\end{equation}
%
Their method has the advantages of remaining on-distribution when
bins become wide and, most importantly, allows the recomputation of
the accumulated effect with different bin-splitting with near-zero
computational cost.

Both approximations ask from the user to blindly decide the bin
density for splitting the feature axis into \(K\) equisized bins, an
approach that is error prone. Setting \(K\) to a small value may hide
fine-grain effects due to large bins while setting \(K\) to a high
value leads to poor bin-effect estimations due to few samples per
bin. In general, the user may receive contradictory explanations for
different \(K\) without a clue for which one to trust.

\section{Uncertainty-Aware ALE (UALE)}
\label{sec:UALE}

UALE extends ALE for quantifying the level of heterogeneous effects
(uncertainty) and automates the bin-splitting. At
Section~\ref{sec:UALE-definition-1} we define UALE, at
Section~\ref{sec:UALE-definition-2} we reformulate it using an
interval-based formulation, at Section~\ref{sec:UALE-approximation} we
provide UALE approximation and at Section~\ref{sec:bin-spliting} we
define and solve the problem of optimal bin-splitting.

\subsection{UALE: ALE With Uncertainty Quantification}
\label{sec:UALE-definition-1}

UALE extends ALE definition of Eq.~(\ref{eq:ALE}) with a component for
quantifying the uncertainty. We denote as \(\mathcal{H}(z)\) the
uncertainty of the local effects at a specific point \(x_s=z\) and we
quantify it with the standard deviation of the local explanations
\(\mathcal{H}(z) = \sigma(z)\), where:

\begin{equation}
  \label{eq:ALE_var}
  \sigma^2(z) = \mathbb{E}_{\Xcb|z}\left [ \left (\dfdx (z, X_c) - \mu(z) \right )^2 \right ] 
\end{equation}
\noindent
The uncertainty emerges from the natural characteristics of the
experiment, i.e.,~the feature correlations existent in the data
generating distribution and the implicit interactions of the black-box
function. We also define the accumulated uncertainty at \(x_s\), as
the accumulation of the standard deviation of the local effects along
the axis:
\begin{equation}
  \label{eq:ALE_acc_unc}
  f^{\mathtt{ALE}}_{\sigma}(x_s) = \int_{x_{s, min}}^{x_s} \sigma(z) \partial z
\end{equation}
\noindent
UALE formulates the effect at a specific point \(x_s\) with a tuple
that consists of the average effect and the uncertainty
\((\mu(z), \sigma(z))\) and visualizes it with a continuous curve and
a confidence region
\(f^{\mathtt{UALE}}(x_s) := f^{\mathtt{ALE}}_{\mu}(x_s) \pm
f^{\mathtt{ALE}}_{\sigma}(x_s)\).

\subsection{Interval-Based Formulation and Approximation}
\label{sec:UALE-definition-2}

In real ML scenarios, all estimations are based on the instances of
the training set. Therefore, we cannot estimate \(\mu(z), \sigma(z)\)
at the granularity of a point, because the probability of observing a
sample inside the region \([x_s - h, x_s + h]\) tends to zero, when
\(h \to 0\). We define the regional-effect \(\mu(z_1, z_2)\) and the
regional-uncertainty \(\mathcal{H}(z_1, z_2) = \sigma(z_1, z_2)\), as
the summary statistics that characterize the effect and the
uncertainty inside the interval \([z_1, z_2)\):

\begin{equation}
  \label{eq:mu_bin}
    \mu(z_1, z_2) = \mathbb{E}_{z \sim \mathcal{U}(z_1,z_2)} [\mu(z)]
    = \frac{\int_{z_1}^{z_2} \mu(z) \partial z}{z_2 - z_1}
\end{equation}

\begin{equation}
  \label{eq:var_bin}
  \sigma^2(z_1, z_2) = \mathbb{E}_{z \sim \mathcal{U}(z_1,z_2)} [\sigma^2(z)] =  \frac{\int_{z_1}^{z_2} \sigma^2(z)  \partial z}{z_2 - z_1}
\end{equation}

%
Intuitively, the regional-effect and the regional-uncertainty quantify
the expected average effect and the expected uncertainty if we
randomly draw a point \(z^*\) given a uniform distribution
\(z^* \sim \mathcal{U}(z_1, z_2)\). If we also define as
\(\mathcal{Z}\) the sequence of \(K+1\) points that partition the
domain of the \(s\)-th feature into \(K\) variable-size intervals,
i.e.  \(\mathcal{Z} = \{z_0, \ldots, z_K\}\), we can redefine UALE
using an interval-based formulation
\(\tilde{f}^{\mathtt{UALE}}_{\mathcal{Z}}(x_s):= \tilde{f}^{\mathtt{ALE}}_{\mathcal{Z}, \mu}(x_s)
\pm \tilde{f}^{\mathtt{ALE}}_{\mathcal{Z}, \sigma}(x_s)\), where:

\begin{equation}
  \label{eq:ALE_2}
  \tilde{f}^{\mathtt{ALE}}_{\mathcal{Z}, \mu}(x_s) = \sum_{k=1}^{k_x} \mu(z_{k-1}, z_k) (z_k - z_{k-1})
\end{equation}

\begin{equation}
  \label{eq:ALE_accumulated_var}
  \tilde{f}^{\mathtt{ALE}}_{\mathcal{Z}, \sigma}(x_s) =  \sum_{k=1}^{k_x} \sigma(z_{k-1}, z_k) (z_k - z_{k-1})
\end{equation}
%

\subsection{Interval-Based Approximation}
\label{sec:UALE-approximation}

For approximating the mean effect (Eq.~\eqref{eq:mu_bin}) and the
uncertainty (Eq.~\eqref{eq:var_bin}) we use the set \(\mathcal{S}_k\)
of dataset instances with the \(s\)-th feature lying inside the
interval
\( \mathcal{S}_k= \{ \mathbf{x}^i : z_{k-1} \leq x^i_s < z_k \} \):

\begin{equation}
  \label{eq:2}
  \hat{\mu}(z_{k-1}, z_k) = \frac{1}{|\mathcal{S}_k|}
  \sum_{i:\mathbf{x}^i \in \mathcal{S}_k} \left [ \dfdx(\mathbf{x}^i)
  \right ]
\end{equation}
%
\begin{equation}
  \label{eq:3}
  \hat{\sigma}^2(z_{k-1}, z_k) = \frac{1}{|\mathcal{S}_k|}
\sum_{i:\mathbf{x}^i \in \mathcal{S}_k} \left ( \dfdx(\mathbf{x}^i) -
  \hat{\mu}(z_1, z_2) \right )^2
\end{equation}

Eq.\eqref{eq:2} is an unbiased estimator of Eq.~(\ref{eq:mu_bin})
under the assumption that the points are uniformly distributed inside
the intervals. The same does not hold in the general case for
Eq.\eqref{eq:3}. In the general case, \(\hat{\sigma}^2(z_{k-1}, z_k)\)
is an unbiased estimator of the \textit{observable} variance
\(\sigma^2_{obs}(z_1, z_2) = \frac{\int_{z_1}^{z_2}
    \mathbb{E}_{\xc|x_s=z} \left [ (f^s(z, \xc) - \mu(z_1, z_2) )^2
    \right] \partial z}{z_2 - z_1}\), which as we prove in Theorem 1
is an overestimation of \(\sigma^2(z_1, z_2)\).

\paragraph{Theorem 1.}
\label{sec:theorem-1}

\textit{If we define the residual \(\rho(z)\) as the difference
  between the expected effect at \(x_s\) and the regional effect, i.e
  \(\rho(z) = \mu(z) - \mu(z_1, z_2)\), then the regional variance
  \(\sigma^2(z_1, z_2)\) equals to:}

\begin{equation}
    \label{eq:bin-uncertainty-proof}
 \sigma_{obs}^2(z_1, z_2) = \sigma^2(z_1, z_2) + \frac{\int_{z_1}^{z_2}\rho^2(z) \partial z}{z_2 - z_1}
\end{equation}
\noindent
Theorem 1 reveals an important connection between the ground-truth and
the observable uncertainty inside a region. For convenience, we denote
as
\(\mathcal{E}^2(z_1, z_2) = \frac{\int_{z_1}^{z_2}\rho^2(z) \partial
  z}{z_2 - z_1}\) the error term that models the expected square
residual inside the interval and as
\(\mathcal{H}_{obs}(z_1, z_2) := \sigma_{obs}(z_1, z_2)\) the
observable uncertainty. It holds that the observable uncertainty is an
overestimation of the correct uncertainty

\begin{equation}
  \label{eq:uncertainty-bin}
  \mathcal{H}^2_{obs}(z_1, z_2) = \mathcal{H}^2(z_1, z_2) + \mathcal{E}^2(z_1, z_2)
\end{equation}
%
and, therefore, the estimation is unbiased only in case
\(\mathcal{E}^2(z_1, z_2) = 0\).

\subsection{Bin-Splitting: Finding Regions With Homogeneous Effects}
\label{sec:bin-spliting}

In this section we formulate bin-splitting as an unsupervised
clustering problem, searching for a solution that compromises two
contradictory objectives. On the one hand, we want to minimize the
error term \(\mathcal{E}\) for approaching an unbiased estimation and
on the other hand, we want wide-enough bins for including a fair
population of samples for a robust estimation of
\(\hat{\mu}(z_1, z_2), \hat{\sigma}(z_1, z_2)\). We set-up the
following optimization problem:

\begin{equation}
  \label{eq:opt}
\begin{aligned}
  \min_{ \mathcal{Z} = \{z_0, \ldots, z_K\}} \quad & \mathcal{L} = \sum_{k=1}^K \tau_k \hat{\sigma}^2(z_{k-1}, z_k) \Delta z_k \\
  \textrm{where} \quad & \Delta z_k = z_k - z_{k-1} \\
  & \tau_k = 1 - \alpha \frac{|S_k|}{N} \\
  \textrm{s.t.} \quad & |\mathcal{S}_k| \geq N_{\mathtt{NPB}}\\
                                     & z_0 = x_{s,min}\\
                                     & z_K = x_{s, max}
\end{aligned}
\end{equation}
%
We search for the sequence of intervals \(z_0, \ldots, z_K\) that
minimizes the sum of the bin costs. The cost of each bin is the
approximated variance \(\hat{\sigma}^2_k\) scaled by the bin length
\(\Delta z_k\) and discounted by the term \(\tau_k\). The term
\(\tau_K\) favors the selection of a bigger bin in case it has similar
variance with the aggregate variance of many smaller ones. The
constraint of at least \(N_{\mathtt{PPB}}\) points per bin sets the
lowest-limit for a \textit{robust} estimation. The user can choose to
what extent they favor the creation of wide bins through the discount
parameter by setting the parameter \(\alpha\) and where they set the
minimum for robust approximation with the parameter
\(N_{\mathtt{PPB}}\). For providing a rough idea, in our experiments
we set \(\alpha = 0.2\) which means that the discount ranges between
\([0\%, 20\%]\) and \(N_{\mathtt{PPB}} = \frac{N}{20}\). It is also
important to clarify that by minimizing
\(\hat{\sigma}^2_k \approx \mathcal{H}_{obs}^2\), we actually minimize
the squared error term \(\mathcal{E}^2\), since the term
\(\mathcal{H}^2_{true}\) is independent of the bin splitting.

\subsubsection{Solve Bin-Splitting with Dynamic Programming}
\label{sec:dynamic-programing}

\todo{Review all the description}
For achieving a computationally-grounded solution we set a threshold
\(K_{max}\) on the maximum number of bins which also discretizes the
solution space. The width of the bin can take discrete values that are
multiple of the minimum step
\(u = \frac{x_{s, max} - x_{s, min}}{K_{max}}\). For defining the
solution, we use two indexes. The index
\(i \in \{0, \ldots, K_{max}\}\) denotes the point \((z_i)\) and the
index \(j \in \{0, \ldots, K_{max}\} \) denotes the position of the
\(j\)-th multiple of the minimum step, i.e., 
\(x_j = x_{s,min} + j \cdot u\). The recursive cost function
\(T(i,j)\) is the cost of setting \(z_i=x_j\):
\begin{equation}
  \label{eq:recursive_cost}
  \mathcal{T}(i,j) = \mathrm{min}_{l \in \{0, \ldots, K_{max}\}} \left [ \mathcal{T}(i-1, l) + \mathcal{B}(x_l, x_j) \right ]
\end{equation}
%
where \(\mathcal{T}(0,j)\) equals zero if \(j=0\) and \(\infty\) in
any other case. \(\mathcal{B}(x_l, x_j)\) denotes the cost of creating a bin
with limits \([x_l, x_j)]\):

\begin{equation}
  \label{eq:cost_step}
  \mathcal{B}(x_l, x_j) = \begin{cases}
                            \infty, & \text{if $x_j > x_l$ or \(|\mathcal{S}_{(x_j, x_l)}| < N\)}\\
                            0, & \text{if $x_j = x_l$}\\
                            \hat{\sigma}^2(x_j, x_l), &\text{if $x_j \leq x_l$}
  \end{cases}
\end{equation}

The optimal solution is given by solving
\(\mathcal{L} = \mathcal{T}(K_{max}, K_{max})\) and keeping track of the sequence of
steps.

\noindent

\begin{itemize}
\item Discuss more aspects (e.g. Computational complexity)
\end{itemize}

\section{SIMULATION EXAMPLES}
\label{sec:simulation-examples}

In XAI it is common~\citep{aas2021explaining, herbinger2022repid} to
use simulation examples for compering methods, because the
data-generating distribution \(p(\mathbf{X})\) and the predictive
function \(f(\cdot)\) are known, and, therefore, the ground-truth is
accesible. We follow this common XAI practice and we split the
evaluation in two groups of examples. The first group
(Section~\ref{sec:simulation-examples-1}) compares UALE definition
against PDP-ICE in modeling the average effect and the
uncertainty. The second group
(Section~\ref{sec:simulation-examples-2}) compares UALE approximation
using the automatic bin-splitting against the fixed-size alternative,
in estimating the average effect and the uncertainty.

\subsection{Case 1: UALE vs PDP-ICE}
\label{sec:simulation-examples-1}

In this simulation, we compare UALE against PDP-ICE. The aim of this
example is to highlight that when features are correlated, PDP and ICE
fail in quantifying the average effect and the heterogeneity, even
when the black-box function is relatatively simple.

We use the following distribution:
\(p(\mathbf{x}) = p(x_3)p(x_2|x_1)p(x_1)\) where
\(x_1 \sim \mathcal{U}(0,1)\), \(x_2 = x_1\) and
\(x_3 \sim \mathcal{N}(0, \sigma_3^2)\) and the following piecewise
linear function:

\begin{equation}
  \label{eq:synth-ex-1-function}
  f(\mathbf{x}) = \begin{cases}
                    f_{\mathtt{lin}} + \alpha f_{\mathtt{int}} & \text{if $f_{\mathtt{lin}} < 0.5$ }\\
                    0.5 - f_{\mathtt{lin}} + \alpha f_{\mathtt{int}} & \text{if $0.5 \leq f_{\mathtt{lin}} < 1$}\\
                    \alpha f_{\mathtt{int}} &\text{otherwise}
                  \end{cases}
\end{equation}
%
where \(f_{\mathtt{lin}}(\mathbf{x}) = a_1 x_1 + a_2 x_2\) and
\(f_{\mathtt{int}}(\mathbf{x}) = x_1x_3\). The linear term
\(f_{\mathtt{lin}}\) includes the two correlated features and the
interaction term \(f_{\mathtt{int}}\) the two non-correlated.  We
evaluate the effect computed by UALE and PDP-ICE in three cases; (a)
without interaction (\(\alpha=0\)) and equal weights (\(a_1=a_2\)),
(b) without interaction (\(\alpha=0\)) and different weigths
(\( a_1 \neq a_2 \)) and (c) with interaction (\(\alpha \neq 0\)) with
equal weights (\(a_1=a_2\)). In all cases, we provide the ground-truth
average effect and uncertainty computed with analytical derivations
(proofs at the Appendix) and we compare it against the approximations
provided by each method, after generating \(N=500\) samples.

\paragraph{No Interaction, Equal weights.}

In this set-up, \(\alpha=0\) (no interaction) and the \(a_1=a_2=1\)
(equal weights). The ground truth effect \(f_\mu^{\mathtt{GT}}(x_1)\)
is: \(x_1\) in \([0, 0.25]\), \(-x_1\) in \([0.25, 0.5]\) and \(0\) in
\([0.5, 1]\). The uncertainty is
\(f^{\mathcal{GT}}_{\sigma^2}(x_1) = 0\) in \([0,1]\) because \(x_1\)
does not interact with ohter variables. In
Figure~\ref{fig:synth-ex-1-case-1}, we observe that PDP effect is
wrong and ICE plots imply the existence of heterogeneous effects. In
contrast, UALE models correctly both the average effect and the zero
uncertainty and approximates it accurately finding a zero aggregate
residual error partitioning of \(x_1\) domain.

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{example_1/dale_feat_0.pdf}
  \includegraphics[width=.23\textwidth]{example_1/pdp_ice_feat_0.pdf}
  \caption{No interaction, Equal weights: Feature effect for \(x_1\)
    using UALE (Left) and PDP-ICE (Right).}
  \label{fig:synth-ex-1-case-1}
\end{figure}

\paragraph{No Interaction, Non-Equal Weights.}

In this set-up, \(\alpha=0\) (no interaction), \(a_1 = 2\) and
\(a_2= 0.5\) (non-equal weights). The non-equal weights have
implications at both the gradient and the interval of the piece-wise
linear regions, i.e., \(f_\mu^{\mathtt{GT}}(x_1)\) is: \(2x_1\) in
\([0, 0.2]\), \(-2x_1\) in \([0.2, 0.4]\) and \(0\) in \([0.4,
1]\). As before, the ground-truth uncertainty is
\(f^{\mathcal{GT}}_{\sigma^2}(x_1) = 0\) because \(x_1\) does not
interact with other features.  In Figure~\ref{fig:ex-synth-1-2}, we
observe that PDP estimation is completely opposite to the ground-truth
effect, i.e.~negative in the region \([0, 0.2)\), positive in
\([0.2, 0.4)\), and the ICE erroneously implies the existence of
heterogeneous effects. As before, ALE quantifies correctly the ground
truth effect, the zero-uncertainty and partioning correctly the
\(x_1\) domain.

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{example_2/dale_feat_0.pdf}
  \includegraphics[width=.23\textwidth]{example_2/pdp_ice_feat_0.pdf}
  \caption{No interaction, Different weights: Feature effect for \(x_1\)
    using UALE (Left) and PDP-ICE (Right).}
  \label{fig:ex-synth-1-2}
\end{figure}

\paragraph{Uncertainty, Equal weights.}

In this set-up, we activate the interaction term, i.e.  \(a=1\),
keeping the weights equal \(a_1=a_2=1\). The interaction term provokes
heterogeneous effects for features \(x_1\) and \(x_3\), because the
local effects at \(x_1\) depend on the unknown value of \(X_3\) and
vice-versa. The effect of \(x_2\) has zero-uncertainty since it does
not appear in any interaction term.  As we observe in
Figure~\ref{fig:ex-synth-1-3}, UALE correcly models the average effect
and the uncertainty in all cases.  On the other hand, PDP-ICE
quantifies correctly the average effect and the uncertainty only in
the case of feature \(x_3\), because \(X_3\) is independent from other
features. For the correlated features (\(x_1, x_2\)) both the average
effect (PDP) and the uncertainty (ICE) are wrong.

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{example_3/dale_feat_0.pdf}
  \includegraphics[width=.23\textwidth]{example_3/pdp_ice_feat_0.pdf}\\
  \includegraphics[width=.23\textwidth]{example_3/dale_feat_1.pdf}
  \includegraphics[width=.23\textwidth]{example_3/pdp_ice_feat_1.pdf}\\
  \includegraphics[width=.23\textwidth]{example_3/dale_feat_2.pdf}
  \includegraphics[width=.23\textwidth]{example_3/pdp_ice_feat_2.pdf}\\
  \caption{With interaction, equal weights: From top to bottom,
    feature effect for features \(\{x_1, x_2, x_3\}\) using UALE (left
    column) and PDP-ICE (right column).}
  \label{fig:ex-synth-1-3}
\end{figure}

\paragraph{Discussion.}

Despite the model's simplicity, PDP-ICE fail in modeling both the
average effect and the uncertainty. The error is due to the method's
formulation, i.e.~the use of the marginal distribution that ignores
correlations between features, and not due to poor approximation
because of limited samples. In contrast, UALE models both correctly
and estimates them accurately extracting the regions with constant
effect.  The examples above do not cover the case of an interaction
term between correlated features, for example a term \(x_1x_2\),
because there is an open debate about the ground-truth effect in this
case~\citep{Gromping2020MAEP}.

\subsection{Case 2: Bin-Splitting}
\label{sec:simulation-examples-2}

In this simulation, we illustrate the advantages of automatic
bin-splitting against the fixed-size alternative. For this reason, we
generate a very big dataset applying dense sampling (\(N=10^6\)) and
we treat the estimation with dense fixed-size bins (\(K=10^3\)) as the
ground-truth UALE. Afterwards, we generate less samples (\(N=500\))
and we compare the fixed-size estimation (for many different \(K\))
against UALE automatic bin-splitting. In all set-ups, we sample from
\(p(\mathbf{x}) = p(x_2|x_1)p(x_1)\) where
\(x_1 \sim \mathcal{U}(0,1)\) and
\(x_2 \sim \mathcal{N}(x_1, \sigma_2^2=0.5)\). We denote as
\(\mathcal{Z^{\mathtt{DP}}} = \{z^{\mathtt{DP}}_{k-1}, \cdots,
z^{\mathtt{DP}}_{K}\}\) the sequence obtained by automatic
bin-splitting and with \(\mathcal{Z^{\mathtt{K}}}\) the fixed-size
splitting with \(K\) bins. The evaluation metrics we report are the
average number of \(t = 30\) independent runs, using each time
\(N=500\) different samples.

\paragraph{Metrics}

The evaluation is done with regard to two metrics counting the mean
error per bin, where the bin error is definde as the absolute
difference of the approximation from the ground-truth. The first
metric, Eq.~\eqref{eq:eval_met_1}, counts the mean bin error wrt
average effect and the second, Eq.~\eqref{eq:eval_met_2} the mean bin
error wrt uncertainty:

\begin{equation}
  \label{eq:eval_met_1}
  \mathcal{L}_{\mathtt{DP|K}}^{\mu} = \frac{1}{|\mathcal{Z}^{\mathtt{DP|K}}|} \sum_{k \in
  \mathcal{Z}^{\mathtt{DP|K}}} | \mu(z_k- z_{k-1}) - \hat{\mu}(z_k-
  z_{k-1}) |
\end{equation}

\begin{equation}
  \label{eq:eval_met_2}
  \mathcal{L}_{\mathtt{DP|K}}^{\sigma} = \frac{1}{|\mathcal{Z}^{\mathtt{DP|K}}|} \sum_{k \in
    \mathcal{Z}^{\mathtt{DP|K}}} | \sigma(z_k- z_{k-1}) - \hat{\sigma}(z_k-
  z_{k-1}) |
\end{equation}

The ground truth \(\mu_k\) (\(\sigma_k\)) is the mean value of the
bin-effects (bin-uncertainties) of the bins inside the interval
defined by the wide bin. For better interpretation of the uncertainty
error, we also provide the mean (per bin) error term
\(\mathcal{L}^{\rho}_{\mathtt{DP|K}} =
\frac{1}{|\mathcal{Z}^{\mathtt{DP|K}}|} \sum_{k \in
  \mathcal{Z}^{\mathtt{DP|K}}} \mathcal{\Epsilon}(z_{k-1}, z_k) \).

\paragraph{Piecewise-Linear Function.}

In this set-up, \(f(\mathbf{x}) = a_1x_1 + x_1x_2\) is a
piecewise-linear function with \(5\) different-width regions, i.e.,
\(a_1\) equals to \(=\{2, -2, 5, -10, 0.5\}\) in the intervals defined
by the sequence \(\{0, 0.2, 0.4, 0.45, 0.5, 1\}\). As we observe, in
Figure~\ref{fig:ex-synth-2-1}, UALE's approximation is better than all
fixed-size alternatives, in both mean effect (\(\mathcal{L}^{\mu}\))
and uncertainty (\(\mathcal{L}^{\sigma}\)). For understanding the
importance of variable-size splitting, we analyze the fixed-bin
error. In the top-right of Figure~\ref{fig:ex-synth-2-1}, we observe
the Law of Large Numbers (LLN); \(\mathcal{L}^{\mu}_{\mathtt{K}}\) and
\(K\) have positive correlation because as \(\Delta x\) becomes
smaller less points contributing to each estimation. The
interpretation of \(\mathcal{L}^{\sigma}\) is more complex. Given that
the smallest piecewise-linear region is \(\Delta x = 0.05 \) any
\(K\) that is not a multiple of \(\frac{1}{\Delta x} = 20\), adds
nuisance uncertainty \(\mathcal{L}^{\rho}_{\mathtt{K}} > 0\) leading
to a biased uncertainty approximation. For these \(K\), uncertainy
error decreases for bigger \(K\) because the length of the bins that
lie in the limit between two piecewise linear regions becomes
smaller. For \(K\) that are multiple of \(20\), i.e.
\(K=\{20, 40, 60, 80, 100\}\), the uncertainty approximation is
unbiased \(\mathcal{L}^{\rho}_{\mathtt{K}} \approx 0\). Here, as in
\(\mathcal{L}^{\mu}\), uncertainty estimation is worse as \(K\) grows
larger (LLN). As we observe in the top left of
Figure~\ref{fig:ex-synth-2-1}, automatic bin-splitting separates the
fine-grain bins, e.g.~regions \([0.4, 0.45]\), \([0.45, 0.5]\), and
unites constant-effect regions into a single bin, e.g.~region
\([0.5, 1]\).

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{example_bin_splitting_1/fig_1.pdf}
  \includegraphics[width=.23\textwidth]{example_bin_splitting_1/fig_2.pdf}\\
  \includegraphics[width=.23\textwidth]{example_bin_splitting_1/fig_3.pdf}
  \includegraphics[width=.23\textwidth]{example_bin_splitting_1/fig_4.pdf}
  \caption{Figure 1}
  \label{fig:ex-synth-2-1}
\end{figure}

\paragraph{Non-Linear Function.}

In this set-up, \(f(\mathbf{x}) = 4x_1^2 + x_2^2 + x_1x_2\), so the
effect is non-linear in \(0 \leq x_1 \leq 1\). In this set-up, wide
bins increase \(\mathcal{L}^{\rho^2}\) making the uncertainty
approximation more biased and narrow bins lead to a worse
approximation. In Figure~\ref{fig:ex-synth-2-2}, we observe that
automatic bin splitting manages to compromise the
conflicting objectives, \todo{Explain better}

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{example_bin_splitting_2/fig_1.pdf}
  \includegraphics[width=.23\textwidth]{example_bin_splitting_2/fig_2.pdf}\\
  \includegraphics[width=.23\textwidth]{example_bin_splitting_2/fig_3.pdf}
  \includegraphics[width=.23\textwidth]{example_bin_splitting_2/fig_4.pdf}
  \caption{Figure 1}
  \label{fig:ex-synth-2-2}
\end{figure}

\section{REAL-WORLD EXAMPLE}

Here, we aim at demonstrating the usefuleness of uncertainty
quantification and the advantages of automatic bin-splitting, on the
real-world California Housing dataset~\citep{pace1997sparse}.

\paragraph{ML setup}

The California Housing is a largely-studied dataset with approximately
\(20000\) training instances, making it appropriate for robust
Monte-Carlo approximations. The dataset contains \(D=8\) numerical
features with characteristics about the building blocks of California,
e.g. latitude, longitude, population of the block or median age of
houses in the block. The target variable is the median value of the
houses inside the block in dollars that ranges between
\([15, 500] \cdots 10^3\), with a mean value of
\(\mu_Y \approx 201 \cdot 10^3 \) and a standard deviation of
\(\sigma_Y \approx 110 \cdot 10^3\).

We exclude instances with missing and outlier values. As outlier we
define the feature values which ar over three standard deviations away
from the mean feature value. We also normalize all features to
zero-mean and unit standard deviation. We split the dataset into
\(N_{tr} = 15639\) training and \(N_{test} = 3910\) test examples
(80/20 split) and we fit a Neural Network with 3 hidden layers of 256,
128 and 36 units respectively. After 15 epochs using the Adam
optimizer with learning rate \(\eta = 0.02\), the model achieves a
normalized mean squared error (R-Squared) of \(0.25\) (0.75), which
corresponds to a MAE of \(37 \cdot 10^3\) dollars.

Below, we illustrate the feature effect for three features: latitude
\(x_2\), population \(x_6\) and median income \(x_8\). The particular
features cover the main FE cases, e.g.~positive/negative trend and
linear/non-linear curve, and they are appropriate for illustration
purposes. The complete results for all features, along with in-depth
information about the preprocessing, training and evaluation parts are
provided in the Appendix.

\paragraph{Uncertainty Quantification}

In real-world datasets, it is infeasible to obtain the ground truth FE
for seamlessly evaluating the competitive methods. We selected the
particular experiment, because, in broad terms, UALE and PDP-ICE plots
aggree in the estimation of the average effect and
uncertainty. Therefore, we can focus on juding the quality of the
information provided by the two methods.

In Figure\ref{fig:ex-real-1}, we observe, from top to bottom, the
effects for the latitude, population and the median income. The effect
of UALE and PDP-ICE are similar for the population and the median
income. The population has a negative impact that progressively
decreases: from 400 to 1500 people the house value decreases with a
rate of \(-150 (\pm 140)\) dollars per added person, a rate that
decreases from \(-80 (\pm 80)\) to \(-60 (\pm 60)\) dollars per added
person as we move from 1500 to 2800 people. The level of uncertainty
indicates significant variance in absolute value of the rate, but in
the grant majority of instances the rate is negative. With the same
inspection, we observe that the median annual income has a positive
impact on the value (all numbers are thousands of dollars):
\(20\pm 15\) per \(10\) of added median income for incomes in
\([8, 15]\), \(32 \pm 20\) per \(10\) added income in \([15, 60]\) and
\(40 \pm 15\) per \(10\) added income in \([60, 70]\). The uncertainty
indicates that there are less heterogeneous effects about the median
income compared to the number of people. In both cases, we can end-up
to the same conclusion by inspecting the PDP-ICE plots. For the
latitude, there is a small difference in the explanations for the
region \([32, 35]\), where UALE estimates a less negative slope with
less uncertainty than PDP, while the explanations are similar for the
range \([35,39]\), where both methods reveal an increase in the
uncertainty around the feature value \(37.5\).

In general, we observe that UALE complements ALE in the quantification
of the heterogeneous effects, similarly to as ICE complements PDP. The
automatic extraction of constant-effect and constant-uncertainty
regions provided by UALE is helpful for an easier interpretation. On
the other hand, ICE plots sometimes are more descritive on locating
the type of heterogeneity, whereas UALE quantifies only the level (not
the type) of heterogeneity.

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_1_ale_auto.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_1_pdp_ice.pdf}\\
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_5_ale_auto.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_5_pdp_ice.pdf}\\
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_7_ale_auto.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_7_pdp_ice.pdf}\\
  \caption{Figure 1}
  \label{fig:ex-real-1}
\end{figure}

\paragraph{Bin Splitting}

In this part, we evaluate the robustness of the approximation using
automatic bin-splitting. Following the evaluation framework of
Section~\ref{sec:simulation-examples-2}, we treat as ground-truth the
effects computed on the full training-set \(N=20000\) with dense
fixed-size bin-splitting (\(K=80\)). Given the big number of samples,
we make the hypothesis that the approximation with dense binning is
close to the ground truth. Afterwards, we randomly select less samples
\(N=1000\) and we compare UALE approximation with all possible
fixed-size alternatives, repeating this process for 30 times for
robust results. In Figure~\ref{fig:ex-real-2}, we illustrate the mean
values for \(\mathcal{L}^{\mu}, \mathcal{L}^{\sigma}\) of the 30
repetitions. We observe that automatic bin-spliting provides (close to
the) best approximation in the three features. In the Appendix, we
provide the same evaluation for all features.

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_mu_err_feature_1.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_var_err_feature_1.pdf}\\
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_mu_err_feature_5.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_var_err_feature_5.pdf}\\
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_mu_err_feature_7.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_var_err_feature_7.pdf}\\
  \caption{Figure 1}
  \label{fig:ex-real-2}
\end{figure}

\section{DISCUSSION}

\todo{Add Conclusion, limitations, and future work}

\subsubsection*{Acknowledgments}
All acknowledgments go at the end of the paper, including thanks to reviewers who gave useful comments, to colleagues who contributed to the ideas, and to funding agencies and corporate sponsors that provided financial support. 
To preserve the anonymity, please include acknowledgments \emph{only} in the camera-ready papers.

\bibliography{biblio}

\end{document}