### 2022-10-12 AISTATS2023/paper.tex
revised method description

           \documentclass[twoside]{article}

\usepackage{aistats2023}
% If your paper is accepted, change the options for the package
% aistats2023 as follows:
%
%\usepackage[accepted]{aistats2023}
%
% This option will print headings for the title of your paper and
% headings for the authors names, plus a copyright note at the end of
% the first column of the first page.

% If you set papersize explicitly, activate the following three lines:
%\special{papersize = 8.5in, 11in}
%\setlength{\pdfpageheight}{11in}
%\setlength{\pdfpagewidth}{8.5in}

% If you use natbib package, activate the following three lines:
\usepackage[round]{natbib}
\renewcommand{\bibname}{References}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}
\bibliographystyle{plainnat}
\usepackage{amsmath}
\usepackage{amssymb}
\DeclareMathOperator*{\argmin}{arg\,min}
\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}

\usepackage{xfrac}
\usepackage{hyperref}
\usepackage{graphicx}
\graphicspath{ {./../examples/} }
% If you use BibTeX in apalike style, activate the following line:
%\bibliographystyle{apalike}

\usepackage{tikz}
\usetikzlibrary{matrix,positioning,arrows.meta,arrows,fit,backgrounds,decorations.pathreplacing}
\tikzset{ mymat/.style={ matrix of math nodes, text height=2.5ex, text
depth=0.75ex, text width=6.00ex, align=center, column
sep=-\pgflinewidth, nodes={minimum height=5.0ex} }, mymats/.style={
mymat, nodes={draw,fill=#1} }, mymat2/.style={ matrix of math nodes,
text height=1.0ex, text depth=0.0ex, minimum width=5ex, % text
width=7.00ex, align=center, column sep=-\pgflinewidth }, }
\usetikzlibrary{shapes.geometric, arrows, backgrounds, scopes}
\usepackage{pgfplots} \pgfplotsset{width=6.75cm, compat=newest}
\usepackage[utf8]{inputenc} \DeclareUnicodeCharacter{2212}{−}
\usepgfplotslibrary{groupplots,dateplot}
\usetikzlibrary{patterns,shapes.arrows}

\newcommand{\dfdx}{f^s}
\newcommand{\xc}{\mathbf{x}_c}
\newcommand{\DY}{\mathbf{\Delta Y}}
\newcommand{\xb}{\mathbf{x}}
\newcommand{\Xcb}{\mathbf{X}_c}
\newcommand{\Xb}{\mathcal{X}}
\newcommand{\todo}[1]{[\textcolor{red}{#1}]}

\begin{document}

% If your paper is accepted and the title of your paper is very long,
% the style will print as headings an error message. Use the following
% command to supply a shorter title of your paper so that it can be
% used as headings.
%
%\runningtitle{I use this title instead because the last one was very long}

% If your paper is accepted and the number of authors is large, the
% style will print as headings an error message. Use the following
% command to supply a shorter version of the authors names so that
% they can be used as headings (for example, use only the surnames)
%
%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}

\twocolumn[

\aistatstitle{Uncertainty-aware Accumulated Local Effects (UALE) for quantifying the heterogeneity of instance-level feature effects}

\aistatsauthor{ Author 1 \And Author 2 \And  Author 3 }

\aistatsaddress{ Institution 1 \And  Institution 2 \And Institution 3 } ]

\begin{abstract}
Accumulated Local Effects (ALE) is a popular explainable AI method that quantifies how a feature influences the decisions of a model, handling well feature correlations. In case of complex interactions between features, instance-level feature effects may deviate from the ALE curve. It is therefore crucial to quantify this deviation, namely, the uncertainty of the effect. In this work, we define Uncertainty-aware ALE (UALE) to quantify and visualize, on a single plot, both the average effect and its uncertainty. We show that UALE quantifies uncertainty effectively, even in case of correlated features. We also note that as in ALE, UALE's approximation requires partitioning the feature domain into non-overlapping intervals (bin-splitting). The average effect and the uncertainty are computed from the instances that lie in each bin. We formally prove that to achieve an unbiased approximation of the uncertainty in each bin, bin-splitting must follow specific constraints. Based on this, we propose a method to determine the optimal intervals, \todo{balancing the estimation bias and variance.} We demonstrate, through synthetic and real datasets, (a) the advantages of modeling the uncertainty with UALE compared to alternative methods and (b) the effectiveness of UALE's appropriate bin splitting for a \todo{good} approximating of the average effect and its uncertainty.
\end{abstract}

\section{INTRODUCTION}

Recently, Machine Learning (ML) has been widely adopted in mission critical domains,
such as healthcare and finance. In such high-stakes areas,  it is important not only to provide accurate predictions but also meaningful explanations. For this reason, there is an increased interest in Explainable
AI (XAI). XAI literature distinguishes between local and global
methods~\citep{Molnar2020interpretable}. Local methods provide
instance-level explanations, i.e., explain the prediction for a
specific instance, whereas global methods summarize the entire model
behavior. Global methods create a global explanation by aggregating the
instance-level explanations into a single interpretable outcome,
usually a number or a plot.

A popular class of global XAI are feature effect (FE)
methods~\citep{Gromping2020MAEP} that quantify the average (over all
instances) effect of a single feature on the output\footnote{FE
  methods also isolate the combined effect of a pair of features to
  the output. Combinations of more than two features are uncommon,
  since they are difficult to estimate and visualize.}. There are
three widely-used FE methods: \emph{Partial Dependence Plots}
(PDP)\citep{friedman2001greedy}, \emph{Marginal Plots}
(MP)\citep{apley2020visualizing} and \emph{Accumulated Local Effects}
(ALE)\citep{apley2020visualizing}. ALE is established as the
state-of-the-art in quantifying the average effect, since PDP and MP
have been criticized~\citep{Gromping2020MAEP} of providing misleading explanations in case of  correlated input features.

When complex interactions between features exist, the instance-level
(local) feature effects may exhibit significant deviation, aka heterogeneity, from the aggregated (global)
explanation, a phenomenon known as aggregation
  bias~\citep{mehrabi2021survey}. We call this deviation \emph{the uncertainty of the global effect}. Aggregation bias leads to
ambiguities; for example, a feature with zero average effect may
indicate (a) no effect on the output or (b) an effect that is highly
positive for some instances and highly negative for some others. As a
result, there is a need for FE methods to quantify the deviation (uncertainty) of
instance-level explanations, in addition to the average effect. In the
case of PDP, such deviation is visualized via the Individual
Conditional Explanations (ICE)~\citep{goldstein2015peeking}. ICE
plots, however, have the same limitations as PDP in case of correlated
features. No method to quantify the deviation of instance-level effects has been proposed for ALE.

% Such ambiguities could be resolved if the heterogeneity of instance-level feature effects is provided in addition to the average effect. 
% In this paper, we measure the heterogeneity as the standard deviation of the instance-level effects and we refer to it using as the \emph{uncertainty of the effect}.

%Therefore, global methods should quantify not only the aggregated (averaged) effect, but also the level of heterogeneity of instance-level explanations.

%There are three widely-used FE methods: \emph{Partial Dependence Plots} (PDP)\citep{friedman2001greedy}, \emph{Marginal Plots}
%(MP)\citep{apley2020visualizing} and \emph{Aggregated Local Effects}
%(ALE)\citep{apley2020visualizing}. ALE has been established as the
%state-of-the-art for measuring the average effect, since PDP and MP
%have been criticized~\citep{Gromping2020MAEP} to compute erroneous
%effects when the input features are (highly) correlated.

In this work, we propose UALE, a global feature-effect method that
extends ALE for modeling the uncertainty of the feature effect.
Our method handles well cases with correlated features. 
As in ALE, we provide an interval-based formulation of UALE. 
This formulation partitions the feature domain into non-overlapping 
intervals (bin-splitting) and defines the average effect (bin-effect) and its uncertainty (bin-uncertainty). We exploit this formulation to provide an UALE approximation. 
As we formally prove, a fixed equi-width partitioning that does not consider
the characteristics of the instance-level effects, which is the case in ALE, 
may lead to an erroneous (biased) estimation of the uncertainty. 
Therefore, we propose a variable-width partitioning that leads to an unbiased approximation.

% %However, ALE faces two limitations, the first concerns ALE definition and the second ALE approximation. Regarding the definition, ALE does not inform the user about the uncertainty. 
% \color{blue}\todo{This should go to related work}In contrast, in PDP the
% heterogeneity of instance-level effects can be identified by exploring
% the Individual Conditional Expectations
% (ICE)~\citep{goldstein2015peeking}. \color{black}
% %Regarding the approximation, ALE requires partitioning the feature domain into non-overlapping intervals (bin-splitting) and estimate the effect inside each interval (bin-effect) from the training set instances that fall inside. 
% \color{blue}\todo{to be moved in related work}
% Finding
% an appropriate sequence of bins is of particular importance, since
% ALE's interpretation is meaningful only inside each
% bin~\citep{molnar2022}. However, this crucial step has not raised the
% appropriate attention. With the current
% approaches~\citep{apley2020visualizing, gkolemis22}, the domain is
% split \emph{blindly}, i.e.~without considering the characteristics of
% the instance-level effects, in \(K\) equisized bins.\color{black}

%In this paper, we present Uncertainty-aware Accumulated Local Effects(UALE), a method that visualizes in a single plot both the av erage feature effect and the uncertainty, i.e.~the level of heterogeneity of instance-level effects. The average effect is defined as in ALE and the uncertainty inherits ALE's advantages, i.e.~takes into account the correlations between features. We also propose an accurate UALE approximation. We formally prove that for an unbiased estimation of the uncertainty, the partitioning of the feature domain must respect specific conditions. Using that, we set-up an optimization problem where we search for a partitioning, a sequence of \emph{variable-size}intervals, (a) with enough instances (accurate estimation) and (b) that respect the above conditions (unbiased estimation). Finally, we propose a method for automatically finding the optimal partitioning. We apply our method in multiple synthetic and real world datasets for evaluating its performance. 
The contributions of this
paper are:
\begin{itemize}
\item A global feature effect method (UALE) that quantifies both the
  average effect and its uncertainty, i.e.,~the deviation of
  instance-level effects.
\item An unbiased approximation of the uncertainty through a dynamic
  variable-size bin partitioning of the feature domain.
%A formal proof of the conditions the feature domain partitioning must fulfill for an unbiased approximation of the uncertainty.
\item The evaluation of UALE on both synthetic and real datasets.
%An accurate UALE approximation by automatically partitioning into variable-size intervals that respect the above conditions.
\end{itemize}

The implementation of our method and the code for reproducing all the
experiments is provided along with the manuscript and will become
publicly available upon acceptance.

\section{BACKGROUND AND RELATED WORK}

In the rest of the paper, we use following notation. Let \(\mathcal{X} \in \mathbb{R}^d\) be the \(d\)-dimensional feature
space, \(\mathcal{Y} \in \mathbb{R}\) the target space and
\(f(\cdot) : \mathcal{X} \rightarrow \mathcal{Y}\) the black-box
function. We use index \(s \in \{1, \ldots, d\}\) for the feature of
interest and \(c = \{1, \ldots, d\} - s\) for the set with all other
indexes. For convenience, we denote the feature vector
\(\xb = (x_1, \cdots , x_s, \cdots, x_D)\) with \((x_s, \xc)\) and the
corresponding random variables
\(X = (X_1, \cdots , X_s, \cdots, X_D)\) with \((X_s, X_c)\). The
training set is \(\mathcal{D} = \{(\xb^i, y^i)\}_{i=1}^N\) sampled
i.i.d. from the distribution \(\mathbb{P}_{X,Y}\). Finally, we use
\(f^{\mathtt{<method>}}(x_s)\) for denoting the \(s\)-th FE, where
\(\mathtt{<method>}\) indicates the particular method in use, for
example ALE. An extensive list with all symbols used in this paper is
provided at Appendix~\ref{sec:list-with-symbols}.

\subsection{Feature Effect Methods}
\label{sec:feat-effect-meth}
The three well-known feature effect methods are: \emph{Partial
  Dependence Plots} (PDP), \emph{Marginal Plots} (MP) and
\emph{Accumulated Local Effects} (ALE).  PDP formulates the global FE
as an expectation over the distribution of \(X_c\), i.e.,
\(f^{\mathtt{PDP}}(x_s) = \mathbb{E}_{X_c}[f(x_s,X_c)]\), whereas MP
as an expectation over the distribution of \(X_c|X_s\), i.e.,
\(f^{\mathtt{MP}}(x_s) = \mathbb{E}_{X_c|x_s}[f(x_s, X_c)]\). Both
methods suffer from misestimations in case of correlated features. PDP
integrates over unrealistic instances and MP computes aggregated
effects, i.e., imputes the combined effect of sets of features to a
single feature~\citep{apley2020visualizing}.

ALE overcomes these problems. Specifically, ALE defines the local
effect of the \(s\)-th feature at a specific point \((z, \xc)\) of the
input space \(\mathcal{X}\) with the partial derivative
\(f^s(z, \xc) = \frac{\partial f}{\partial x_s} (z, \xc)\). All the
local explanations at \(z\) are then weighted by the conditional
distribution \(p(\xc|z)\) and are averaged, to produce the averaged
effect \(\mu(z)\). ALE plot is the accumulation of the averaged local
effects:

\begin{equation}
  \label{eq:ALE}
  f^{\mathtt{ALE}}(x_s) = \int_{x_{s,min}}^{x_s} \underbrace{\mathbb{E}_{X_c|X_s=z}\left [f^s (z, X_c)\right ]}_{\mu(z)} \partial z
\end{equation}
%
where \(x_{s,min}\) is the minimum value of the \(s\)-th feature.

\subsection{Heterogeneity Of Instance-Level Effects}
\label{sec:quant-heter-effects}

The global effect is computed as an expectation over local
(instance-level) effects. In addition to this global effect, it is
important to know to what extent the local effects deviate from the
global explanation.

ICE plots and similar methods (e.g., d-ICE
plots~\citep{goldstein2015peeking}) provide a set of curves
illustrated on top-of PDP. Each curve corresponds to one instance of
the dataset, \(f^{\mathtt{ICE}}_i(x_s) = f(x_s, \xc^i)\). The user
then visually observes if the curves are homogeneous, i.e., all
instances have similar effect plots, and to what extent they deviate
from the PDP. There are methods try to automate the aforementioned
visual exploration, by grouping ICE plots into
clusters~\citep{molnar2020model, herbinger2022repid,
  britton2019vine}. Unfortunately, these methods are subject to the
failure modes of PDPs in cases of correlated
features~\citep{baniecki2021fooling}, as we also confirm in our
experimental analysis (c.f., Section~\ref{sec:simulation-examples-1}).

Other approaches, like H-Statistic \citep{friedman2008predictive},
Greenwell's interaction index \citep{greenwell2018simple} or SHAP
interaction values \citep{lundberg2018consistent} quantify
interactions between the input features. A strong interaction is
  an implicit indication of the existence of heterogeneous effects. These
  methods, however, do not quantify directly the level of heterogeneity.

To the best of our knowledge, no work exist so far that quantifies the
heterogeneous effects (uncertainty) based on the formulation of ALE.

\subsection{ALE Approximation}
\label{sec:ale-approximation}

ALE is estimated from the available dataset instances. \citep{apley2020visualizing} proposed to divide the feature
domain in \(K\) bins of equal size and to estimate the local effects
in each bin by evaluating the model \(f\) at the bin limits:

\begin{equation}
  \label{eq:ALE_accumulated_mean_est}
  \hat{f}^{\mathtt{ALE}}(x_s) = \sum_{k=1}^{k_x} \frac{1}{|\mathcal{S}_k|} \sum_{i:\mathbf{x}^i \in
    \mathcal{S}_k} \left [ f(z_{k}, \xc^i) - f(z_{k-1}, \xc^i)) \right ]
\end{equation}
%
where \(k_x\) the index of the bin that \(x_s\) belongs to,
i.e. \(k_x: z_{k_x-1} \leq x_s < z_{k_x} \) and \(\mathcal{S}_k\) is
the set of training instances of the \(k\)-th bin, i.e.
\( \mathcal{S}_k = \{ \xb^i : z_{k-1} \leq x^i_s < z_{k} \}
\). \citep{gkolemis22} proposed the Differential ALE (DALE) that
computes the local effects on the training instances using
auto-differentiation:
\begin{equation}  \label{eq:DALE_accumulated_mean_est}
  \hat{f}^{\mathtt{DALE}}(x_s) = \Delta x \sum_{k=1}^{k_x} \frac{1}{|\mathcal{S}_k|} \sum_{i:\mathbf{x}^i \in
    \mathcal{S}_k} f^s(\mathbf{x}^i)
\end{equation}
%
Their method has significant computational advantages and allows the
recomputation of the accumulated effect with different bin-splitting
with near-zero additional computational cost. Moreover, by avoiding
the use of artificial samples at the bin limits DALE allows the use of
wider bins without out-of-distribution sampling. \todo{Maybe connect with our work}. In both cases, the
approximations partition the feature domain in \(K\) equally-sized
bins, without considering the underlying local effects.

% as
% input and perform an equally-sized bin splitting of the attribute
% domain. The choice of \(K\) is therefore critical and introduces
% significant limitations. Specifically, setting \(K\) to a small value
% may hide fine-grain effects due to large bins while setting \(K\) to a
% high value leads to poor bin-effect estimations due to few samples per
% bin. In general, the user may receive contradictory explanations for
% different \(K\) values without a clue for which one to trust.

\section{Uncertainty-Aware ALE (UALE)}
\label{sec:UALE}

In this Section~\ref{sec:UALE-definition-1}, we define UALE. Then, we
provide an interval-based formulation that partitions the feature
domain into non-overlapping intervals (bin-splitting) and defines the
bin-effect and the bin-uncertainty
(Section~\ref{sec:UALE-definition-2}). Then, we present the UALE
approximation (Section~\ref{sec:UALE-approximation}), formally proving
the bin-splitting requirements for an unbiased estimation of the
uncertainty. Based on this, we present an optimization method to
determine an appropriate variable-width partitioning
(Section~\ref{sec:bin-spliting}).

\subsection{UALE Definition And Interval-Based Formulation}
\label{sec:UALE-definition-1}

We quantify the uncertainty of the local effects at \(x_s=z\) with the
standard deviation of the local explanations, \(\sigma(z)\), where:

\begin{equation}
  \label{eq:ALE_var}
  \sigma^2(z) = \mathbb{E}_{X_c|X_s=z}\left [ \left (\dfdx (z, X_c) - \mu(z) \right )^2 \right ] 
\end{equation}
\noindent
The uncertainty emerges from (a) the feature correlations and (b) the
implicit feature interactions of the black-box function. We also
define the accumulated uncertainty at \(x_s\), as the accumulation of
the standard deviation of the local effects:

\begin{equation}
  \label{eq:ALE_acc_unc}
  f^{\mathtt{ALE}}_{\sigma}(x_s) = \int_{x_{s, min}}^{x_s} \sigma(z) \partial z
\end{equation}
\noindent

UALE defines the effect as a pair that consists of the average effect
and the uncertainty \((\mu(z), \sigma(z))\). For visualizing it, we
use the accumulated average effect, as defined in Eq.~(\ref{eq:ALE}),
and the accumulated uncertainty, as defined in
Eq.~(\ref{eq:ALE_acc_unc}). Figure~\ref{fig:UALE-figure} illustrates
an example \todo{Describe after putting the figure}.

\begin{figure}
  \centering
  \caption{UALE-concept-figure}
  \label{fig:UALE-figure}
\end{figure}

For the interval-based formulation, we define the bin-effect
\(\mu(z_1, z_2)\) and the corresponding bin-uncertainty
\(\sigma(z_1, z_2)\) as:

\begin{equation}
  \label{eq:mu_bin}
    \mu(z_1, z_2) = \mathbb{E}_{z \sim \mathcal{U}(z_1,z_2)} [\mu(z)]
    = \frac{\int_{z_1}^{z_2} \mu(z) \partial z}{z_2 - z_1}
\end{equation}

\begin{equation}
  \label{eq:var_bin}
  \sigma^2(z_1, z_2) = \mathbb{E}_{z \sim \mathcal{U}(z_1,z_2)} [\sigma^2(z)] =  \frac{\int_{z_1}^{z_2} \sigma^2(z)  \partial z}{z_2 - z_1}
\end{equation}

%
Intuitively, if we randomly draw a point \(z^*\) given a uniform
distribution \(z^* \sim \mathcal{U}(z_1, z_2)\), the bin-effect
(Eq.~(\ref{eq:mu_bin})) and the bin-uncertainty
(Eq.~(\ref{eq:var_bin})) are the expected average effect and the
expected uncertainty, respectively. Also, if we define \(\mathcal{Z}\)
to be the sequence of \(K+1\) points that partition the domain of the
\(s\)-th feature into \(K\) variable-size intervals, i.e.
\(\mathcal{Z} = \{z_0, \ldots, z_K\}\), we can reformulate UALE to:
\begin{equation}
  \label{eq:ALE_2}
  \tilde{f}^{\mathtt{ALE}}_{\mathcal{Z}, \mu}(x_s) = \sum_{k=1}^{k_x} \mu(z_{k-1}, z_k) (z_k - z_{k-1})
\end{equation}

\begin{equation}
  \label{eq:ALE_accumulated_var}
  \tilde{f}^{\mathtt{ALE}}_{\mathcal{Z}, \sigma}(x_s) =  \sum_{k=1}^{k_x} \sigma(z_{k-1}, z_k) (z_k - z_{k-1})
\end{equation}
%
where \(k_x\) is the index of the bin such that
\( z_{k_x - 1} \leq x_s < z_{k_x}\).

\subsection{UALE Interval-Based Approximation}
~\label{sec:UALE-approximation}

For approximating the bin-effect and the bin-uncertainty, we use the
set \(\mathcal{S}_k\) of dataset instances with the \(s\)-th feature
lying inside the \(k\)-th bin, i.e.,
\( \mathcal{S}_k= \{ \mathbf{x}^i : z_{k-1} \leq x^i_s < z_k \}
\). The bin-effect is approximated with,

\begin{equation}
  \label{eq:mu_bin_approx}
  \hat{\mu}(z_{k-1}, z_k) = \frac{1}{|\mathcal{S}_k|}
  \sum_{i:\mathbf{x}^i \in \mathcal{S}_k} \left [ \dfdx(\mathbf{x}^i)
  \right ]
\end{equation}

%
which, as shown by\citep{gkolemis22}, is an unbiased estimator of
Eq.~(\ref{eq:mu_bin}) under the assumption that the points are
uniformly distributed inside the intervals. The approximation for the
uncertainty (Eq.~(\ref{eq:var_bin})) from the available instances is
defined as follows:

%
\begin{equation}
  \label{eq:var_bin_approx}
  \hat{\sigma}^2(z_{k-1}, z_k) = \frac{1}{|\mathcal{S}_k|}
\sum_{i:\mathbf{x}^i \in \mathcal{S}_k} \left ( \dfdx(\mathbf{x}^i) -
  \hat{\mu}(z_1, z_2) \right )^2
\end{equation}
%
In the Appendix~\ref{sec:proof-1}, we show that
\(\hat{\sigma}^2(z_{k-1}, z_k)\) is an unbiased estimator of
\(\sigma^2_*(z_1, z_2) = \frac{\int_{z_1}^{z_2} \mathbb{E}_{\xc|x_s=z}
  \left [ (f^s(z, \xc) - \mu(z_1, z_2) )^2 \right] \partial z}{z_2 -
  z_1} \)
%
which, in the general case, overestimates
\(\sigma^2(z_1, z_2)\), as shown in Theorem~\ref{sec:theorem-1}.

\begin{theorem}
  \label{sec:theorem-1}
  If we define (a) the residual \(\rho(z)\) as the difference between
  the expected effect at \(x_s\) and the regional effect, i.e
  \(\rho(z) = \mu(z) - \mu(z_1, z_2)\), and (b) \(\mathcal{E}(z_1, z_2)\) as the
  mean squared residual of the bin, i.e.
  \(\mathcal{E}(z_1, z_2) = \frac{\int_{z_1}^{z_2}\rho^2(z) \partial
    z}{z_2 - z_1}\), then it holds
\begin{equation}
    \label{eq:bin-uncertainty-proof}
 \sigma_*^2(z_1, z_2) = \sigma^2(z_1, z_2) + \mathcal{E}^2(z_1, z_2)
\end{equation}
  \end{theorem}

\begin{proof}
The proof is at Appendix~\ref{sec:proof-of-3-1}.
\end{proof}

\noindent

Based on Theorem\ref{sec:theorem-1}, the estimation is unbiased only
when \(\mathcal{E}^2(z_1, z_2) = 0\).

\subsection{Bin-Splitting}
\label{sec:bin-spliting}

The quality of interval-based approximation is affected by (a) the
population of samples in each bin and (b) the error term
\(\mathcal{E}(z_1, z_2)\) of each bin. On the one hand, we favor a
partitioning of the feature domain in large-bins to allow a robust
estimation of \(\hat{\mu}(z_1, z_2), \hat{\sigma}(z_1, z_2)\) from a
sufficient population of samples. On the other hand, we should
minimize the aggregation of bin errors, i.e.,
\( \mathcal{E}^2_{\mathcal{Z}} = \sum_{k=1}^K\mathcal{E}^2(z_1, z_2)
\Delta z_k\), where \(\mathcal{Z} = \{z_0, \cdots, z_K\}\) and
\(\Delta z_k = z_k - z_{k-1}\). For convenience, we refer to
\(\mathcal{E}^2_{\mathcal{Z}}\) as aggregated error. We search for a
partitioning \(\mathcal{Z}_*\) that balances this trade-off.

Due to the Corollary~\ref{sec:coroll}, minimizing
\( \mathcal{E}^2_{\mathcal{Z}} \) is equivalent to minimizing
\(\sigma_*^2(z_{k-1}, z_k)\Delta z_k\), which can be directly
estimated by \(\hat{\sigma}^2(z_{k-1}, z_k) \Delta z_k \).

\begin{corollary}
  \label{sec:coroll}
  The minimization of the aggregation of bin variances is equivalent
  to the minimization of the aggregated error:
  \begin{equation}
  \begin{gathered}
    \label{eq:2}
    \argmin_{ \mathcal{Z} = \{z_0, \ldots, z_K\}} \mathcal{E}^2_{\mathcal{Z}}  = \argmin_{ \mathcal{Z} = \{z_0, \ldots, z_K\}} \sum_{k=1}^K\sigma_*^2(z_1, z_2) \Delta z_k 
  \end{gathered}
  \end{equation}
\end{corollary}

\begin{proof}
~\label{sec:coroll-1}
  The proof is at Appendix~\ref{sec:proof-of-coroll}
  \end{proof}

Based on the above, we set-up the following optimization problem:

\begin{equation}
  \label{eq:opt}
\begin{aligned}
  \min_{ \mathcal{Z} = \{z_0, \ldots, z_K\}} \quad & \mathcal{L} = \sum_{k=1}^K \tau_k \hat{\sigma}^2(z_{k-1}, z_k) \Delta z_k \\
  \textrm{where} \quad & \Delta z_k = z_k - z_{k-1} \\
  & \tau_k = 1 - \alpha \frac{|S_k|}{N} \\
  \textrm{s.t.} \quad & |\mathcal{S}_k| \geq N_{\mathtt{NPB}}\\
                                     & z_0 = x_{s,min}\\
                                     & z_K = x_{s, max}
\end{aligned}
\end{equation}
%
The objective \(\mathcal{L}\) searches for the partitioning
\(\mathcal{Z}_* = \{ z_0, \ldots, z_K \} \) that minimizes the
aggregated error \(\mathcal{E}_{\mathcal{Z}}^2\), while the term
\(\tau_K\) favors the partitioning that contains bigger bins, in case
many partitionings have similar aggregated error. The constraint of at
least \(N_{\mathtt{PPB}}\) points per bin sets the lowest-limit for a
\textit{robust} estimation. The user can choose to what extent they
favor the creation of wide bins through (a) the parameter \(\alpha\)
that controls the discount \(\tau_k\) and (b) the parameter
\(N_{\mathtt{PPB}}\) that sets the minimum population per bin. For
providing a rough idea, in our experiments we set \(\alpha = 0.2\)
which means that the discount ranges between \([0\%, 20\%]\) and
\(N_{\mathtt{PPB}} = \frac{N}{20}\).

\subsubsection{Solve Bin-Splitting with Dynamic Programming}
\label{sec:dynamic-programing}

\todo{Review all the description}
For achieving a computationally-grounded solution we set a threshold
\(K_{max}\) on the maximum number of bins which also discretizes the
solution space. The width of the bin can take discrete values that are
multiple of the minimum step
\(u = \frac{x_{s, max} - x_{s, min}}{K_{max}}\). For defining the
solution, we use two indexes. The index
\(i \in \{0, \ldots, K_{max}\}\) denotes the point \((z_i)\) and the
index \(j \in \{0, \ldots, K_{max}\} \) denotes the position of the
\(j\)-th multiple of the minimum step, i.e., 
\(x_j = x_{s,min} + j \cdot u\). The recursive cost function
\(T(i,j)\) is the cost of setting \(z_i=x_j\):
\begin{equation}
  \label{eq:recursive_cost}
  \mathcal{T}(i,j) = \mathrm{min}_{l \in \{0, \ldots, K_{max}\}} \left [ \mathcal{T}(i-1, l) + \mathcal{B}(x_l, x_j) \right ]
\end{equation}
%
where \(\mathcal{T}(0,j)\) equals zero if \(j=0\) and \(\infty\) in
any other case. \(\mathcal{B}(x_l, x_j)\) denotes the cost of creating a bin
with limits \([x_l, x_j)]\):

\begin{equation}
  \label{eq:cost_step}
  \mathcal{B}(x_l, x_j) = \begin{cases}
                            \infty, & \text{if $x_j > x_l$ or \(|\mathcal{S}_{(x_j, x_l)}| < N\)}\\
                            0, & \text{if $x_j = x_l$}\\
                            \hat{\sigma}^2(x_j, x_l), &\text{if $x_j \leq x_l$}
  \end{cases}
\end{equation}

The optimal solution is given by solving
\(\mathcal{L} = \mathcal{T}(K_{max}, K_{max})\) and keeping track of the sequence of
steps.

\noindent

\begin{itemize}
\item Discuss more aspects (e.g. Computational complexity)
\end{itemize}

\section{SIMULATION EXAMPLES}
\label{sec:simulation-examples}

In XAI it is common~\citep{aas2021explaining, herbinger2022repid} to
use simulation examples for compering methods, because the
data-generating distribution \(p(\mathbf{X})\) and the predictive
function \(f(\cdot)\) are known, and, therefore, the ground-truth is
accesible. We follow this common XAI practice and we split the
evaluation in two groups of examples. The first group
(Section~\ref{sec:simulation-examples-1}) compares UALE definition
against PDP-ICE in modeling the average effect and the
uncertainty. The second group
(Section~\ref{sec:simulation-examples-2}) compares UALE approximation
using the automatic bin-splitting against the fixed-size alternative,
in estimating the average effect and the uncertainty.

\subsection{Case 1: UALE vs PDP-ICE}
\label{sec:simulation-examples-1}

In this simulation, we compare UALE against PDP-ICE. The aim of this
example is to highlight that when features are correlated, PDP and ICE
fail in quantifying the average effect and the heterogeneity, even
when the black-box function is relatatively simple.

We use the following distribution:
\(p(\mathbf{x}) = p(x_3)p(x_2|x_1)p(x_1)\) where
\(x_1 \sim \mathcal{U}(0,1)\), \(x_2 = x_1\) and
\(x_3 \sim \mathcal{N}(0, \sigma_3^2)\) and the following piecewise
linear function:

\begin{equation}
  \label{eq:synth-ex-1-function}
  f(\mathbf{x}) = \begin{cases}
                    f_{\mathtt{lin}} + \alpha f_{\mathtt{int}} & \text{if $f_{\mathtt{lin}} < 0.5$ }\\
                    0.5 - f_{\mathtt{lin}} + \alpha f_{\mathtt{int}} & \text{if $0.5 \leq f_{\mathtt{lin}} < 1$}\\
                    \alpha f_{\mathtt{int}} &\text{otherwise}
                  \end{cases}
\end{equation}
%
where \(f_{\mathtt{lin}}(\mathbf{x}) = a_1 x_1 + a_2 x_2\) and
\(f_{\mathtt{int}}(\mathbf{x}) = x_1x_3\). The linear term
\(f_{\mathtt{lin}}\) includes the two correlated features and the
interaction term \(f_{\mathtt{int}}\) the two non-correlated.  We
evaluate the effect computed by UALE and PDP-ICE in three cases; (a)
without interaction (\(\alpha=0\)) and equal weights (\(a_1=a_2\)),
(b) without interaction (\(\alpha=0\)) and different weigths
(\( a_1 \neq a_2 \)) and (c) with interaction (\(\alpha \neq 0\)) with
equal weights (\(a_1=a_2\)). In all cases, we provide the ground-truth
average effect and uncertainty computed with analytical derivations
(proofs at the Appendix) and we compare it against the approximations
provided by each method, after generating \(N=500\) samples.

\paragraph{No Interaction, Equal weights.}

In this set-up, \(\alpha=0\) (no interaction) and the \(a_1=a_2=1\)
(equal weights). The ground truth effect \(f_\mu^{\mathtt{GT}}(x_1)\)
is: \(x_1\) in \([0, 0.25]\), \(-x_1\) in \([0.25, 0.5]\) and \(0\) in
\([0.5, 1]\). The uncertainty is
\(f^{\mathcal{GT}}_{\sigma^2}(x_1) = 0\) in \([0,1]\) because \(x_1\)
does not interact with ohter variables. In
Figure~\ref{fig:synth-ex-1-case-1}, we observe that PDP effect is
wrong and ICE plots imply the existence of heterogeneous effects. In
contrast, UALE models correctly both the average effect and the zero
uncertainty and approximates it accurately finding a zero aggregate
residual error partitioning of \(x_1\) domain.

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{example_1/dale_feat_0.pdf}
  \includegraphics[width=.23\textwidth]{example_1/pdp_ice_feat_0.pdf}
  \caption{No interaction, Equal weights: Feature effect for \(x_1\)
    using UALE (Left) and PDP-ICE (Right).}
  \label{fig:synth-ex-1-case-1}
\end{figure}

\paragraph{No Interaction, Non-Equal Weights.}

In this set-up, \(\alpha=0\) (no interaction), \(a_1 = 2\) and
\(a_2= 0.5\) (non-equal weights). The non-equal weights have
implications at both the gradient and the interval of the piece-wise
linear regions, i.e., \(f_\mu^{\mathtt{GT}}(x_1)\) is: \(2x_1\) in
\([0, 0.2]\), \(-2x_1\) in \([0.2, 0.4]\) and \(0\) in \([0.4,
1]\). As before, the ground-truth uncertainty is
\(f^{\mathcal{GT}}_{\sigma^2}(x_1) = 0\) because \(x_1\) does not
interact with other features.  In Figure~\ref{fig:ex-synth-1-2}, we
observe that PDP estimation is completely opposite to the ground-truth
effect, i.e.~negative in the region \([0, 0.2)\), positive in
\([0.2, 0.4)\), and the ICE erroneously implies the existence of
heterogeneous effects. As before, ALE quantifies correctly the ground
truth effect, the zero-uncertainty and partioning correctly the
\(x_1\) domain.

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{example_2/dale_feat_0.pdf}
  \includegraphics[width=.23\textwidth]{example_2/pdp_ice_feat_0.pdf}
  \caption{No interaction, Different weights: Feature effect for \(x_1\)
    using UALE (Left) and PDP-ICE (Right).}
  \label{fig:ex-synth-1-2}
\end{figure}

\paragraph{Uncertainty, Equal weights.}

In this set-up, we activate the interaction term, i.e.  \(a=1\),
keeping the weights equal \(a_1=a_2=1\). The interaction term provokes
heterogeneous effects for features \(x_1\) and \(x_3\), because the
local effects at \(x_1\) depend on the unknown value of \(X_3\) and
vice-versa. The effect of \(x_2\) has zero-uncertainty since it does
not appear in any interaction term.  As we observe in
Figure~\ref{fig:ex-synth-1-3}, UALE correcly models the average effect
and the uncertainty in all cases.  On the other hand, PDP-ICE
quantifies correctly the average effect and the uncertainty only in
the case of feature \(x_3\), because \(X_3\) is independent from other
features. For the correlated features (\(x_1, x_2\)) both the average
effect (PDP) and the uncertainty (ICE) are wrong.

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{example_3/dale_feat_0.pdf}
  \includegraphics[width=.23\textwidth]{example_3/pdp_ice_feat_0.pdf}\\
  \includegraphics[width=.23\textwidth]{example_3/dale_feat_1.pdf}
  \includegraphics[width=.23\textwidth]{example_3/pdp_ice_feat_1.pdf}\\
  \includegraphics[width=.23\textwidth]{example_3/dale_feat_2.pdf}
  \includegraphics[width=.23\textwidth]{example_3/pdp_ice_feat_2.pdf}\\
  \caption{With interaction, equal weights: From top to bottom,
    feature effect for features \(\{x_1, x_2, x_3\}\) using UALE (left
    column) and PDP-ICE (right column).}
  \label{fig:ex-synth-1-3}
\end{figure}

\paragraph{Discussion.}

Despite the model's simplicity, PDP-ICE fail in modeling both the
average effect and the uncertainty. The error is due to the method's
formulation, i.e.~the use of the marginal distribution that ignores
correlations between features, and not due to poor approximation
because of limited samples. In contrast, UALE models both correctly
and estimates them accurately extracting the regions with constant
effect.  The examples above do not cover the case of an interaction
term between correlated features, for example a term \(x_1x_2\),
because there is an open debate about the ground-truth effect in this
case~\citep{Gromping2020MAEP}.

\subsection{Case 2: Bin-Splitting}
\label{sec:simulation-examples-2}

In this simulation, we illustrate the advantages of automatic
bin-splitting against the fixed-size alternative. For this reason, we
generate a very big dataset applying dense sampling (\(N=10^6\)) and
we treat the estimation with dense fixed-size bins (\(K=10^3\)) as the
ground-truth UALE. Afterwards, we generate less samples (\(N=500\))
and we compare the fixed-size estimation (for many different \(K\))
against UALE automatic bin-splitting. In all set-ups, we sample from
\(p(\mathbf{x}) = p(x_2|x_1)p(x_1)\) where
\(x_1 \sim \mathcal{U}(0,1)\) and
\(x_2 \sim \mathcal{N}(x_1, \sigma_2^2=0.5)\). We denote as
\(\mathcal{Z^{\mathtt{DP}}} = \{z^{\mathtt{DP}}_{k-1}, \cdots,
z^{\mathtt{DP}}_{K}\}\) the sequence obtained by automatic
bin-splitting and with \(\mathcal{Z^{\mathtt{K}}}\) the fixed-size
splitting with \(K\) bins. The evaluation metrics we report are the
average number of \(t = 30\) independent runs, using each time
\(N=500\) different samples.

\paragraph{Metrics}

The evaluation is done with regard to two metrics counting the mean
error per bin, where the bin error is definde as the absolute
difference of the approximation from the ground-truth. The first
metric, Eq.~\eqref{eq:eval_met_1}, counts the mean bin error wrt
average effect and the second, Eq.~\eqref{eq:eval_met_2} the mean bin
error wrt uncertainty:

\begin{equation}
  \label{eq:eval_met_1}
  \mathcal{L}_{\mathtt{DP|K}}^{\mu} = \frac{1}{|\mathcal{Z}^{\mathtt{DP|K}}|} \sum_{k \in
  \mathcal{Z}^{\mathtt{DP|K}}} | \mu(z_k- z_{k-1}) - \hat{\mu}(z_k-
  z_{k-1}) |
\end{equation}

\begin{equation}
  \label{eq:eval_met_2}
  \mathcal{L}_{\mathtt{DP|K}}^{\sigma} = \frac{1}{|\mathcal{Z}^{\mathtt{DP|K}}|} \sum_{k \in
    \mathcal{Z}^{\mathtt{DP|K}}} | \sigma(z_k- z_{k-1}) - \hat{\sigma}(z_k-
  z_{k-1}) |
\end{equation}

The ground truth \(\mu_k\) (\(\sigma_k\)) is the mean value of the
bin-effects (bin-uncertainties) of the bins inside the interval
defined by the wide bin. For better interpretation of the uncertainty
error, we also provide the mean (per bin) error term
\(\mathcal{L}^{\rho}_{\mathtt{DP|K}} =
\frac{1}{|\mathcal{Z}^{\mathtt{DP|K}}|} \sum_{k \in \mathcal{Z}^{\mathtt{DP|K}}} \mathcal{E}(z_{k-1}, z_k) \).

\paragraph{Piecewise-Linear Function.}

In this set-up, \(f(\mathbf{x}) = a_1x_1 + x_1x_2\) is a
piecewise-linear function with \(5\) different-width regions, i.e.,
\(a_1\) equals to \(=\{2, -2, 5, -10, 0.5\}\) in the intervals defined
by the sequence \(\{0, 0.2, 0.4, 0.45, 0.5, 1\}\). As we observe, in
Figure~\ref{fig:ex-synth-2-1}, UALE's approximation is better than all
fixed-size alternatives, in both mean effect (\(\mathcal{L}^{\mu}\))
and uncertainty (\(\mathcal{L}^{\sigma}\)). For understanding the
importance of variable-size splitting, we analyze the fixed-bin
error. In the top-right of Figure~\ref{fig:ex-synth-2-1}, we observe
the Law of Large Numbers (LLN); \(\mathcal{L}^{\mu}_{\mathtt{K}}\) and
\(K\) have positive correlation because as \(\Delta x\) becomes
smaller less points contributing to each estimation. The
interpretation of \(\mathcal{L}^{\sigma}\) is more complex. Given that
the smallest piecewise-linear region is \(\Delta x = 0.05 \) any
\(K\) that is not a multiple of \(\frac{1}{\Delta x} = 20\), adds
nuisance uncertainty \(\mathcal{L}^{\rho}_{\mathtt{K}} > 0\) leading
to a biased uncertainty approximation. For these \(K\), uncertainy
error decreases for bigger \(K\) because the length of the bins that
lie in the limit between two piecewise linear regions becomes
smaller. For \(K\) that are multiple of \(20\), i.e.
\(K=\{20, 40, 60, 80, 100\}\), the uncertainty approximation is
unbiased \(\mathcal{L}^{\rho}_{\mathtt{K}} \approx 0\). Here, as in
\(\mathcal{L}^{\mu}\), uncertainty estimation is worse as \(K\) grows
larger (LLN). As we observe in the top left of
Figure~\ref{fig:ex-synth-2-1}, automatic bin-splitting separates the
fine-grain bins, e.g.~regions \([0.4, 0.45]\), \([0.45, 0.5]\), and
unites constant-effect regions into a single bin, e.g.~region
\([0.5, 1]\).

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{example_bin_splitting_1/fig_1.pdf}
  \includegraphics[width=.23\textwidth]{example_bin_splitting_1/fig_2.pdf}\\
  \includegraphics[width=.23\textwidth]{example_bin_splitting_1/fig_3.pdf}
  \includegraphics[width=.23\textwidth]{example_bin_splitting_1/fig_4.pdf}
  \caption{Figure 1}
  \label{fig:ex-synth-2-1}
\end{figure}

\paragraph{Non-Linear Function.}

In this set-up, \(f(\mathbf{x}) = 4x_1^2 + x_2^2 + x_1x_2\), so the
effect is non-linear in \(0 \leq x_1 \leq 1\). In this set-up, wide
bins increase \(\mathcal{L}^{\rho^2}\) making the uncertainty
approximation more biased and narrow bins lead to a worse
approximation. In Figure~\ref{fig:ex-synth-2-2}, we observe that
automatic bin splitting manages to compromise the
conflicting objectives, \todo{Explain better}

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{example_bin_splitting_2/fig_1.pdf}
  \includegraphics[width=.23\textwidth]{example_bin_splitting_2/fig_2.pdf}\\
  \includegraphics[width=.23\textwidth]{example_bin_splitting_2/fig_3.pdf}
  \includegraphics[width=.23\textwidth]{example_bin_splitting_2/fig_4.pdf}
  \caption{Figure 1}
  \label{fig:ex-synth-2-2}
\end{figure}

\section{REAL-WORLD EXAMPLE}

Here, we aim at demonstrating the usefuleness of uncertainty
quantification and the advantages of automatic bin-splitting, on the
real-world California Housing dataset~\citep{pace1997sparse}.

\paragraph{ML setup}

The California Housing is a largely-studied dataset with approximately
\(20000\) training instances, making it appropriate for robust
Monte-Carlo approximations. The dataset contains \(D=8\) numerical
features with characteristics about the building blocks of California,
e.g. latitude, longitude, population of the block or median age of
houses in the block. The target variable is the median value of the
houses inside the block in dollars that ranges between
\([15, 500] \cdots 10^3\), with a mean value of
\(\mu_Y \approx 201 \cdot 10^3 \) and a standard deviation of
\(\sigma_Y \approx 110 \cdot 10^3\).

We exclude instances with missing and outlier values. As outlier we
define the feature values which ar over three standard deviations away
from the mean feature value. We also normalize all features to
zero-mean and unit standard deviation. We split the dataset into
\(N_{tr} = 15639\) training and \(N_{test} = 3910\) test examples
(80/20 split) and we fit a Neural Network with 3 hidden layers of 256,
128 and 36 units respectively. After 15 epochs using the Adam
optimizer with learning rate \(\eta = 0.02\), the model achieves a
normalized mean squared error (R-Squared) of \(0.25\) (0.75), which
corresponds to a MAE of \(37 \cdot 10^3\) dollars.

Below, we illustrate the feature effect for three features: latitude
\(x_2\), population \(x_6\) and median income \(x_8\). The particular
features cover the main FE cases, e.g.~positive/negative trend and
linear/non-linear curve, and they are appropriate for illustration
purposes. The complete results for all features, along with in-depth
information about the preprocessing, training and evaluation parts are
provided in the Appendix.

\paragraph{Uncertainty Quantification}

In real-world datasets, it is infeasible to obtain the ground truth FE
for seamlessly evaluating the competitive methods. We selected the
particular experiment, because, in broad terms, UALE and PDP-ICE plots
aggree in the estimation of the average effect and
uncertainty. Therefore, we can focus on juding the quality of the
information provided by the two methods.

In Figure\ref{fig:ex-real-1}, we observe, from top to bottom, the
effects for the latitude, population and the median income. The effect
of UALE and PDP-ICE are similar for the population and the median
income. The population has a negative impact that progressively
decreases: from 400 to 1500 people the house value decreases with a
rate of \(-150 (\pm 140)\) dollars per added person, a rate that
decreases from \(-80 (\pm 80)\) to \(-60 (\pm 60)\) dollars per added
person as we move from 1500 to 2800 people. The level of uncertainty
indicates significant variance in absolute value of the rate, but in
the grant majority of instances the rate is negative. With the same
inspection, we observe that the median annual income has a positive
impact on the value (all numbers are thousands of dollars):
\(20\pm 15\) per \(10\) of added median income for incomes in
\([8, 15]\), \(32 \pm 20\) per \(10\) added income in \([15, 60]\) and
\(40 \pm 15\) per \(10\) added income in \([60, 70]\). The uncertainty
indicates that there are less heterogeneous effects about the median
income compared to the number of people. In both cases, we can end-up
to the same conclusion by inspecting the PDP-ICE plots. For the
latitude, there is a small difference in the explanations for the
region \([32, 35]\), where UALE estimates a less negative slope with
less uncertainty than PDP, while the explanations are similar for the
range \([35,39]\), where both methods reveal an increase in the
uncertainty around the feature value \(37.5\).

In general, we observe that UALE complements ALE in the quantification
of the heterogeneous effects, similarly to as ICE complements PDP. The
automatic extraction of constant-effect and constant-uncertainty
regions provided by UALE is helpful for an easier interpretation. On
the other hand, ICE plots sometimes are more descritive on locating
the type of heterogeneity, whereas UALE quantifies only the level (not
the type) of heterogeneity.

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_1_ale_auto.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_1_pdp_ice.pdf}\\
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_5_ale_auto.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_5_pdp_ice.pdf}\\
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_7_ale_auto.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_7_pdp_ice.pdf}\\
  \caption{Figure 1}
  \label{fig:ex-real-1}
\end{figure}

\paragraph{Bin Splitting}

In this part, we evaluate the robustness of the approximation using
automatic bin-splitting. Following the evaluation framework of
Section~\ref{sec:simulation-examples-2}, we treat as ground-truth the
effects computed on the full training-set \(N=20000\) with dense
fixed-size bin-splitting (\(K=80\)). Given the big number of samples,
we make the hypothesis that the approximation with dense binning is
close to the ground truth. Afterwards, we randomly select less samples
\(N=1000\) and we compare UALE approximation with all possible
fixed-size alternatives, repeating this process for 30 times for
robust results. In Figure~\ref{fig:ex-real-2}, we illustrate the mean
values for \(\mathcal{L}^{\mu}, \mathcal{L}^{\sigma}\) of the 30
repetitions. We observe that automatic bin-spliting provides (close to
the) best approximation in the three features. In the Appendix, we
provide the same evaluation for all features.

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_mu_err_feature_1.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_var_err_feature_1.pdf}\\
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_mu_err_feature_5.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_var_err_feature_5.pdf}\\
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_mu_err_feature_7.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_var_err_feature_7.pdf}\\
  \caption{Figure 1}
  \label{fig:ex-real-2}
\end{figure}

\section{DISCUSSION}

\todo{Add Conclusion, limitations, and future work}

\subsubsection*{Acknowledgments}
All acknowledgments go at the end of the paper, including thanks to reviewers who gave useful comments, to colleagues who contributed to the ideas, and to funding agencies and corporate sponsors that provided financial support. 
To preserve the anonymity, please include acknowledgments \emph{only} in the camera-ready papers.

\bibliography{biblio}

\appendix

\section{List with Symbols}
\label{sec:list-with-symbols}

\section{Show that is unbiased estimator}
\label{sec:proof-1}

\section{Proof of Theorem~\ref{sec:theorem-1}}
\label{sec:proof-of-3-1}

\section{Proof of Corollary~\ref{sec:coroll-1}}
\label{sec:proof-of-coroll}

\end{document}